---
title: "Bulk RNA-Seq DGE Pipeline"
author: "Pedro Lopez"
date: "260302"
output: html_document
editor_options: 
  chunk_output_type: inline
---


#Input relevant info
```{r}
# Initials <- readline(prompt = "Enter your initials: ")
# 
# cell_line <- readline(prompt = "Enter cell line: ")
# 
# perturbation <- readline(prompt = "Enter perturbation of cell line: ")
# 
# Seq_date <- readline(prompt = "Enter date of RNAseq: ")
# 
# Replicates <- readline(prompt = "Enter number of biological replicates: ")
# 
# dir_name <- paste0(Initials, "_",cell_line, "_", perturbation)
# 
# BAM_files <- rstudioapi::selectDirectory(caption = "select folder with all BAM files", path = "/media/bree")
# 
# gtf.path <- rstudioapi::selectFile(caption = "select .gtf file used for alignment", path = "/media/bree")
# 
# org = readline(prompt = "what organism is being analyzed?: ")
# 
# counts_runner <- as.logical(readline(prompt = "Do you want to generate read count matrix? (T or F) "))
# 
# if(counts_runner == F){
#   counts_matrix <- rstudioapi::selectFile(caption = "select counts matrix excel file")
# }
# 
# rMATS_runner <- as.logical(readline(prompt = "Do you want to run rMATS? (T or F) "))
# 
# if(rMATS_runner == F){
#   MATS.path <- rstudioapi::selectDirectory(caption = "select rMATS output folder from previous run")
# }
# 
# 
# Start = Sys.time()
# 
# gtf <- read.delim(
#   gtf.path,
#   header = FALSE,
#   comment.char = "#",
#   stringsAsFactors = FALSE,
# )
# 
# End = Sys.time()
# print(End - Start)
# colnames(gtf) <- c(
#   "chr","source","feature","start","end",
#   "score","strand","frame","attribute"
# )
# 

```

#create file directories
```{r}
# outputdir <- rstudioapi::selectDirectory(caption = "Select folder for path generation", path = "/media/bree")

outputdir_base <- normalizePath(file.path(outputdir, paste0("RNASeqAnalysis_",dir_name)), mustWork = FALSE)
if(!dir.exists(outputdir_base)){
dir.create(outputdir_base, recursive = TRUE, showWarnings = FALSE)
}
outputdir_SEQ <- normalizePath(file.path(outputdir_base, paste0("DESeq2_",dir_name)), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ)){
dir.create(outputdir_SEQ, recursive = TRUE, showWarnings = FALSE)
}

outputdir_SEQ_allsashimi <- normalizePath(file.path(outputdir_SEQ, "allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ_allsashimi)){
dir.create(outputdir_SEQ_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_SEQ_GO <- normalizePath(file.path(outputdir_SEQ, "GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ_GO)){
dir.create(outputdir_SEQ_GO, recursive = TRUE, showWarnings = FALSE)
}

outputdir_SEQ_GSEA <- normalizePath(file.path(outputdir_SEQ, "GSEA_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ_GSEA)){
dir.create(outputdir_SEQ_GSEA, recursive = TRUE, showWarnings = FALSE)
}

outputdir_SEQ_GSEA_SignP <- normalizePath(file.path(outputdir_SEQ_GSEA, "Signed_Pvalue"), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ_GSEA_SignP)){
dir.create(outputdir_SEQ_GSEA_SignP, recursive = TRUE, showWarnings = FALSE)
}

outputdir_SEQ_GSEA_STAT <- normalizePath(file.path(outputdir_SEQ_GSEA, "STAT_rnk"), mustWork = FALSE)
if(!dir.exists(outputdir_SEQ_GSEA_STAT)){
dir.create(outputdir_SEQ_GSEA_STAT, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS <-normalizePath(file.path(outputdir_base, paste0("rMATS_", dir_name)), mustWork = FALSE)
if(!dir.exists(outputdir_MATS)){
dir.create(outputdir_MATS, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATSfiles <- normalizePath(file.path(outputdir_MATS, "rMATS_output"), mustWork = FALSE)
if(!dir.exists(outputdir_MATSfiles)){
dir.create(outputdir_MATSfiles, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_sashimi <- normalizePath(file.path(outputdir_MATS, "sashimi_plots"), mustWork = F)
if(!dir.exists(outputdir_MATS_sashimi)){
  dir.create(outputdir_MATS_sashimi, recursive = T, showWarnings = F)
}

outputdir_MATS_SE_allsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "SE_allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_SE_allsashimi)){
dir.create(outputdir_MATS_SE_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_MXE_allsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "MXE_allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_MXE_allsashimi)){
dir.create(outputdir_MATS_MXE_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_RI_allsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "RI_allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_RI_allsashimi)){
dir.create(outputdir_MATS_RI_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A3SS_allsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "A3SS_allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A3SS_allsashimi)){
dir.create(outputdir_MATS_A3SS_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A5SS_allsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "A5SS_allsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A5SS_allsashimi)){
dir.create(outputdir_MATS_A5SS_allsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_SE_EXPsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "SE_EXPsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_SE_EXPsashimi)){
dir.create(outputdir_MATS_SE_EXPsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_MXE_EXPsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "MXE_EXPsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_MXE_EXPsashimi)){
dir.create(outputdir_MATS_MXE_EXPsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_RI_EXPsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "RI_EXPsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_RI_EXPsashimi)){
dir.create(outputdir_MATS_RI_EXPsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A3SS_EXPsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "A3SS_EXPsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A3SS_EXPsashimi)){
dir.create(outputdir_MATS_A3SS_EXPsashimi, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A5SS_EXPsashimi <- normalizePath(file.path(outputdir_MATS_sashimi, "A5SS_EXPsashimi"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A5SS_EXPsashimi)){
dir.create(outputdir_MATS_A5SS_EXPsashimi, recursive = TRUE, showWarnings = FALSE)
}

#make folder for all MATS GO plots
outputdir_MATS_GO <- normalizePath(file.path(outputdir_MATS, "GO_plots"), mustWork = F) 
if(!dir.exists(outputdir_MATS_GO)){
  dir.create(outputdir_MATS_GO, recursive = T, showWarnings = F)
}

outputdir_MATS_SE_GO <- normalizePath(file.path(outputdir_MATS_GO, "SE_GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_SE_GO)){
dir.create(outputdir_MATS_SE_GO, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_MXE_GO <- normalizePath(file.path(outputdir_MATS_GO, "MXE_GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_MXE_GO)){
dir.create(outputdir_MATS_MXE_GO, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_RI_GO <- normalizePath(file.path(outputdir_MATS_GO, "RI_GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_RI_GO)){
dir.create(outputdir_MATS_RI_GO, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A3SS_GO <- normalizePath(file.path(outputdir_MATS_GO, "A3SS_GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A3SS_GO)){
dir.create(outputdir_MATS_A3SS_GO, recursive = TRUE, showWarnings = FALSE)
}

outputdir_MATS_A5SS_GO <- normalizePath(file.path(outputdir_MATS_GO, "A5SS_GO_plots"), mustWork = FALSE)
if(!dir.exists(outputdir_MATS_A5SS_GO)){
dir.create(outputdir_MATS_A5SS_GO, recursive = TRUE, showWarnings = FALSE)
}


```

```{r}
bam_files <- list.files(BAM_files , full.names = TRUE)
bam_files <- gsub("\\\\", "/", bam_files)



gg.bams <- c(unlist(bam_files[which(grepl("\\.bam$", bam_files)&grepl(cell_line, bam_files))]))

#make .txt files for Ctrl and KD bam paths
ctrl_bams <- c(gg.bams[which(!grepl(perturbation, basename(gg.bams)))]) %>%
  sapply(function(x){
    bam_name <- str_split_i(x, "/", -1)
    paste0("/bams/", bam_name)
  }) %>%
  paste(collapse = ",") %>%
  writeLines(con = file.path(normalizePath(outputdir_MATS), "ctrl_bams.txt"))
ctrl_bams.path <- file.path(normalizePath(outputdir_MATS), "ctrl_bams.txt")
ctrl_bams_name <- str_split_i(ctrl_bams.path, "/", -1)

KD_bams <- c(gg.bams[which(grepl(perturbation, basename(gg.bams)))]) %>%
  sapply(function(x){
    bam_name <- str_split_i(x, "/", -1)
    paste0("/bams/", bam_name)
  }) %>%
  paste(collapse = ",") %>%
  writeLines(con = file.path(normalizePath(outputdir_MATS), "KD_bams.txt"))
KD_bams.path <- file.path(normalizePath(outputdir_MATS), "KD_bams.txt")
KD_bams_name <- str_split_i(KD_bams.path, "/", -1)

```

## generate counts for DESeq2
```{r}
if(counts_runner == T){
#counting call, feed in bam files and annotations, mark as paired end

ctrl_bams2 <- c(gg.bams[which(!grepl(perturbation, basename(gg.bams)))])
KD_bams2 <- c(gg.bams[which(grepl(perturbation, basename(gg.bams)))])
  
counts_mat_new <- featureCounts(files = c(ctrl_bams2, KD_bams2),
                                isGTFAnnotationFile = TRUE,
                                annot.ext = gtf.path, 
                                isPairedEnd = TRUE , strandSpecific = 2, 
                                allowMultiOverlap = T, countMultiMappingReads = T)
countdata <- counts_mat_new$counts %>%as.data.frame()
#fix colnames

colname_old <- colnames(countdata)

colnames(countdata) <- sapply(colname_old, function(x){
  new_name <- str_split_1(x, "_CKD")[[1]]
  paste(new_name, "_", sep = "")
})

genes_vec <- rownames(countdata)
countdata <- mutate(countdata, Gene = genes_vec, .before = colnames(countdata)[1])
}else{
  counts_mat_new <- read_excel(counts_matrix, if(length(excel_sheets(counts_matrix)) > 1){sheet = "counts matrix"} )
  countdata <- counts_mat_new %>% as.data.frame()
}
#copy and paste your spreadsheet into read_excel(___). 
fileName = paste0(cell_line, "_", perturbation, "_DESeq2.xlsx")

outfile = paste(outputdir_SEQ, fileName, sep = "/")

#README sheet at the start of excel
Sheet <- c("Overall description", "", "", "counts matrix", "", "", "DESeq2 output", "", "",
                       "FDR_0.05", "", "", "|log2FoldChange|_1","","","","","","")
t_ <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description <- c("List all output data associated with DESeq2 pipeline", "", "", "raw read counts used for DESeq2 run", "", "", "unfiltered DESeq2 output. Includes log2FoldChange; log2foldchange standard error (lfcSE); p-value and p adj value (padj or FDR); ensembl ID; gene name (SYMBOL); TPM; Description of gene", "", "", "filtered for padjusted values less than 0.05", "", "", "filtered FDR_0.05 filtered data set further to include only genes with an absolute log2FoldChange greater than 1", "", "", Initials, cell_line, perturbation, date())
README_df <- data.frame(Sheet = Sheet, t_ = t_, Sheet_description = Sheet_description )

wb <- createWorkbook()

if("README" %in% names(wb)){
  print("updating current worksheet")
}else{
addWorksheet(wb, "README")
}
writeData(wb, "README", README_df)
if("counts matrix" %in% names(wb)){
  print("updating current worksheet")
}else{
addWorksheet(wb, "counts matrix")
}
writeData(wb, "counts matrix", countdata)
saveWorkbook(wb, file = outfile, overwrite = TRUE)

#preliminary filtering for counts >= 20 (KD or Ctrl)

SEQ_cutoff <- 20

# SEQ_ctrl <- countdata[,2:(ncol(countdata)-1)/2] %>% rowmeans()
# SEQ_KD <- countdata[,(ncol(countdata)-1)/2:ncol(countdata)] %>% sapply(function(x){mean(as.integer(x))})


SEQ_ctrl <- rowMeans(countdata[,2:(1 + ((ncol(countdata)-1)/2))])

SEQ_KD <- rowMeans(countdata[,(2 + ((ncol(countdata)-1)/2)):ncol(countdata)] )


countdata2 <- countdata[which(SEQ_ctrl >= SEQ_cutoff | SEQ_KD >= SEQ_cutoff),]
if("counts_>=20" %in% names(wb)){
  print("updating current sheet")
}else{
addWorksheet(wb, "counts_>=20")
}
writeData(wb, "counts_>=20", countdata2)
saveWorkbook(wb, file = outfile, overwrite = T)
```

## Prepare data for Analysis with DESeq2
```{r}
## Upload raw read count matrix as a table

counts_mat <- read_excel(outfile, sheet = "counts_>=20")

## Convert the table to a dataframe

counts_mat<-as.data.frame(counts_mat)

## Assign the gene symbol/ID to row names, remove the gene symbol/ID column, 
# asummes that there is a column called Gene
row.names(counts_mat) <- counts_mat$Gene
counts_mat <- counts_mat[ ,-which(colnames(counts_mat) %in% 'Gene')]


## Assign condition and other variables (Ensure the order of the factor matches
## the position/column of the sample within the raw read count matrix)

Group <- factor(c(rep("Ctrl",length(ctrl_bams2)), rep("KD",length(KD_bams2))))
# Batch <- factor(c("Batch1","Batch2","Batch1","Batch2")) <- removed due to RNASeq protocol mod to remove                                                               batch effect 260310


## Create the colData variable (table of sample information) for DESeq2 
## based on the conditions 
coldata<- data.frame(row.names=colnames(counts_mat),Group)
```

## Differential Gene Expression Analysis with DESeq2
```{r}
## Generate a DESEQDataSet File from the matrix, column data, and group variables

dds <- DESeqDataSetFromMatrix(countData = counts_mat,
                              colData = coldata,
                              design = ~Group)

## Run the DESeq function

dds <- DESeq(dds)

## Generate Dispersion Plot for QC purposes

Disp_path <- file.path(normalizePath(outputdir_SEQ), "qc-dispersions.pdf")
pdf(Disp_path, width = 7, height=5)
plotDispEsts(dds, main="Dispersion plot")
dev.off()

## Perform the differential gene expression analysis and acquire the results 
## between the two conditions (use ashr method to generate log2FoldChange 
## values)

res <- results(dds, alpha = 0.05)
res <- lfcShrink(dds, contrast = c("Group","KD","Ctrl"),
type = "ashr")


## Generate a MA plot (scatterplot of the log fold change and mean average of
## normalized counts)

MA_path <- file.path(normalizePath(outputdir_SEQ), "MAplot.png")
png(MA_path, 1000, 1000, pointsize=20)
plotMA(res,ylim=c(-5,5))
dev.off()


## Generate volcano plot

"Gene" <- names(res)[1]
volcanoplot <- function (res, lfcthresh=1, sigthresh=0.05, main="Volcano Plot",
legendpos="bottomright", labelsig=TRUE, textcx=1, ...)
{
with(res, plot(log2FoldChange, -log10(pvalue), pch=20, main=main, ...))
with(subset(res, padj<sigthresh ), points(log2FoldChange, -log10(pvalue),
pch=20, col="red", ...))
with(subset(res, abs(log2FoldChange)>lfcthresh), points(log2FoldChange,
-log10(pvalue), pch=20, col="orange", ...))
with(subset(res, padj<sigthresh & abs(log2FoldChange)>lfcthresh),
points(log2FoldChange, -log10(pvalue), pch=20, col="green", ...))
if (labelsig) {
require(calibrate)
4
with(subset(res, padj<sigthresh & abs(log2FoldChange)>lfcthresh),
textxy(log2FoldChange, -log10(pvalue), labs=NULL, cex=textcx, ...))
}
legend(legendpos, xjust=1, yjust=1, legend=c(paste("FDR<",sigthresh,sep=""),
paste("|LogFC|>",lfcthresh,sep=""), "both"),
pch=20, col=c("red","orange","green"))
}
vol_path <- file.path(normalizePath(outputdir_SEQ), "volcano.png")
png(vol_path, 1200, 1000, pointsize=20)
volcano <- volcanoplot(res, lfcthresh=1, sigthresh=0.05, textcx=.8)
dev.off()

## Perform a variance stabilizing transformation on the DESeq data and create matrix
## with vst transformed count data

vstdata <- vst(dds)
vstMatrix <- assay(vstdata)

##Generate Sample-Sample Distance Heatmap with vst() function

sampleDists <- as.dist(1-cor(vstMatrix))
sampleDistMatrix <- as.matrix(sampleDists)
annoCol <- as.data.frame(colData(dds))
annoCol=annoCol[c(1)]

heat_path <- file.path(normalizePath(outputdir_SEQ), "heatmap.png")
heat_map <- pheatmap(sampleDistMatrix,
clustering_distance_rows=sampleDists,
clustering_distance_cols=sampleDists ,annotation_col = annoCol,
filename = heat_path,
width = 8,
height = 10)
heat_map

```

## OPTIONAL Alternative Sample Distance Plot
```{r}
# Transpose count matrix and calculate distances using dist()
sampleDists <- assay(vstdata) %>% t() %>% dist()

# Convert distance dataframe to matrix
sampleDistMatrix <- as.matrix(sampleDists)

# Choose continuous palette from RColorBrewer
colors <- colorRampPalette(rev(brewer.pal(9, "Blues")))(255)

# Plot sample distance heatmap with ComplexHeatmap
altHeat_path <- file.path(normalizePath(outputdir_SEQ), "Alt_sampleDistance_plot.png")
pheatmap(sampleDistMatrix, cluster_rows = T, cluster_columns = T,
        col = colors,main= cell_line,
        filename = altHeat_path)
```


## Generate PCAplot
```{r}
library(ggplot2)
# rld = rlog(dds)
PCA <- plotPCA(vstdata, intgroup = "Group", ntop = 2000)
pca_path <- file.path(normalizePath(outputdir_SEQ), "PCA_plot.png")
# ggsave(pca_path, plot = PCA, width = 8, height = 6, dpi = 300)
ggsave(pca_path, plot = PCA, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
PCA
```


```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
library(org.Mm.eg.db)
```

```{r}
#modify dds to contain gene symbol, ENSEMBL and ENTREZ ids
res$ENSEMBL = gsub("\\..*","",row.names(res))
 

```

```{r}
#add factor for switch between human and mouse genome
  #do tolower for org
if(org == "human"){
  org_genome = org.Hs.eg.db
}else if( org == "mouse"){
  org_genome = org.Mm.eg.db
}else{
  print("ERROR")
}

res$gene = mapIds(org_genome,
                    keys = res$ENSEMBL,
                    column = "SYMBOL",
                    keytype = "ENSEMBL",
                    multiVals = "first")

res$ENTREZ = mapIds(org_genome,
                    keys = res$ENSEMBL,
                    column = "ENTREZID",
                    keytype = "ENSEMBL",
                    multiVals = "first")
head(res)
```

## Generate log2 normalized gene expression scatterplot 
```{r}
## generate dataframe of normalized expression values for a gene of interest

 KD_gene_row <- which(res$gene %in% perturbation) 

 KDexpression <- plotCounts(dds = dds ,gene = rownames(res)[KD_gene_row],intgroup = "Group",returnData = TRUE)

 ##create a new dataframe and convert the normalized counts to log2 normalized
 ##counts for visualization purposes

 logKDexpression <-KDexpression
 logKDexpression$count<-log2(logKDexpression$count)

##create the plot
 l2f <- ggplot(logKDexpression , aes(x=Group, y=count)) + ylab(paste0("Log2 Normalized ",  perturbation," Expression")) + xlab("Group") +  geom_point(shape=1) + ggtitle(cell_line) + theme(plot.title = element_text(hjust = 0.5))+geom_text(aes(label = rownames(logKDexpression)),hjust = -0.2, vjust = 0.5)
 l2f_path <- file.path(normalizePath(outputdir_SEQ), "l2f_norm_gene_expresScatterplot.png")
 # ggsave(l2f_path, plot = l2f, width = 8, height = 6, dpi = 300)
 ggsave(l2f_path, plot = l2f, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
 l2f
```


## Generate Heatmap of all Significant DE Genes with p-value <0.05 and
## L2FC < -1 (Sig. Downregulated Genes in KD)
```{r}
sigChanges <- rownames(res)[res$padj < 0.05 & res$log2FoldChange < -1 ]
sigMat <- vstMatrix[rownames(vstMatrix) %in% sigChanges,]
sigHDWN <- pheatmap(sigMat, scale="row", show_rownames = FALSE,show_colnames = FALSE,
col=greenred(75), border_color = NA, annotation_col = annoCol)
sigHeatDWN_path <- file.path(normalizePath(outputdir_SEQ), "HeatsigGene_down.png")
# ggsave(sigHeatDWN_path, plot = sigHDWN, width = 8, height = 6, dpi = 300)
ggsave(sigHeatDWN_path, plot = sigHDWN, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
sigHDWN
```

## Generate Heatmap of all Significant DE Genes with p-value <0.05 and
## L2FC > 1 (Sig. Upregulated Genes in KD)
```{r}
sigChanges <- rownames(res)[res$padj < 0.05 & res$log2FoldChange > 1 ]
sigMat <- vstMatrix[rownames(vstMatrix) %in% sigChanges,]
sigHUP <- pheatmap(sigMat, scale="row", show_rownames = FALSE,show_colnames = FALSE,
col=greenred(75), border_color = NA, annotation_col = annoCol)
sigHeatUP_path <- file.path(normalizePath(outputdir_SEQ), "HeatsigGene_up.png")
# ggsave(sigHeatUP_path, plot = sigHUP, width = 8, height = 6, dpi = 300)
ggsave(sigHeatUP_path, plot = sigHUP, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
sigHUP
```
#print out list of genes that are siginificantly different from Ctrl to KD



#Generate TPM data frame from counts matrix
```{r}
rownames(counts_mat) <- gsub("\\..*", "", rownames(counts_mat))
if(org == "human"){
with_TPM  <-IOBR::count2tpm(counts_mat, idType = "Ensembl", org = "hsa")
}else if(org == "mouse"){
  with_TPM  <-IOBR::count2tpm(counts_mat, idType = "Ensembl", org = "mmus")
}
with_TPM$gene <- rownames(with_TPM)



res_df <- as.data.frame(res)
rownames(res_df) <- gsub("\\..*", "", rownames(res_df))


library(mygene)

res_df<- res_df[!is.na(res_df$padj), ] 
res_df <- res_df[order(res_df$padj),]


res_geneList <- res_df$gene

res_df <- cbind(
  res_df,
  with_TPM[match(res_df$gene, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
res_df <- res_df %>%
  mutate(Gene = res_df$gene, .before = 1)
res_df <- res_df %>% 
  mutate(gene = res_df$gene)

res_df <- res_df %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("DESeq2 output" %in% names(wb)){
  print("updating current worksheet")
}else{
addWorksheet(wb, "DESeq2 output")
}
writeData(wb, "DESeq2 output", res_df)
saveWorkbook(wb, file = outfile, overwrite = TRUE)
```

```{r}
#make sheet with filters for cutoffs of DESEq2 output
res_df_padj <- res_df %>%
  filter(padj < 0.05)
if("FDR_0.05" %in% names(wb)){
  print("updatding current worksheet")
}else{
addWorksheet(wb, "FDR_0.05")
}
writeData(wb, "FDR_0.05", res_df_padj)
saveWorkbook(wb, file = outfile, overwrite = TRUE)

res_df_allfilt <- res_df_padj %>%
  filter(abs(log2FoldChange) >= 1 )

gene_info <- getGenes(geneids = res_df_allfilt$ENSEMBL, fields = "summary") 
gene_deduped <- gene_info[which(!duplicated(gene_info$query)),]
gene_desc <- as.vector(gene_deduped$summary)

res_df_allfilt <- res_df_allfilt%>%
  mutate(Description = gene_desc)

if("|log2FoldChange| _ 1" %in% names(wb)){
  print("updating current worksheet")
}else{
addWorksheet(wb, "|log2FoldChange| _ 1")
}
writeData(wb, "|log2FoldChange| _ 1", res_df_allfilt)
saveWorkbook(wb, file = outfile, overwrite = TRUE)
```

##DESeq GO Analysis
```{r}
library(clusterProfiler)
library(WebGestaltR)
```

```{r}

sigUp <- subset(res_df, padj < 0.05 & log2FoldChange >=1)
sigUp <- sigUp[!is.na(sigUp$ENSEMBL),]
sigUp <- sigUp[!is.na(sigUp$log2FoldChange),]
sigUp <- sigUp[!is.na(sigUp$padj),]
genes_sig_Up <- sigUp$ENSEMBL 

write(genes_sig_Up, file = paste0(outputdir_SEQ, "/",cell_line, "_gene_sigUP.txt"))

sigDown <- subset(res_df, padj < 0.05 & log2FoldChange <= -1 )
sigDown <- sigDown[!is.na(sigDown$ENSEMBL),]
sigDown <- sigDown[!is.na(sigDown$log2FoldChange),]
sigDown <- sigDown[!is.na(sigDown$padj),]
genes_sig_down <- sigDown$ENSEMBL

write(genes_sig_down, file = paste0(outputdir_SEQ, "/",cell_line, "_gene_sigDWN.txt"))

sigAll <- res_df[!is.na(res_df$ENSEMBL),]
sigAll <- sigAll$ENSEMBL

write(sigAll, file = paste0(outputdir_SEQ, "/",cell_line, "_gene_sigBKG.txt"))

wb_GOredund = createWorkbook()

library(org.Mm.eg.db)

if(org == "human"){
  id_org = org.Hs.eg.db
}else if(org == "mouse"){
  id_org = org.Mm.eg.db
}



```

#GO noredundant
```{r}
#set up WebGestaltR function for BP
if(org == "human"){
no_web_call_BP <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){
#set if for organism type                                                                                                                     
  WebGestaltR(
    enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
no_web_call_MF <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
no_web_call_KEGG <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
no_web_call_Hall <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

no_web_call_Panther <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_Panther'){
    WebGestaltR(
      enrichMethod = "ORA",
      organism = "hsapiens",
      enrichDatabase = path,
      
      interestGene = x,
      interestGeneType = id,
      
      referenceGene = bkg,
      referenceGeneType = id,
      
      minNum = 5,
      maxNum = 500,
      sigMethod = "fdr",
      fdrMethod = "BH",
      fdrThr = cutoff,
      topThr = 10000,
      
      saveRawGseaResult = FALSE,
      isOutput = FALSE
    )
} 

no_web_call_Reactome <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_Reactome'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

}else if(org == "mouse"){
  no_web_call_BP <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){
 #set if for organism type                                                                                                                     
  WebGestaltR(
    enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
no_web_call_MF <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
no_web_call_KEGG <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
no_web_call_Hall <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
} 

#WebGestaltR for Panther
no_web_call_Panther <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_Panther'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR for Panther
no_web_call_Reactome <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'pathway_Reactome'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

}

```


```{r}
#GO analysis for BP (UP and DOWN)
# wb_DeGO = createWorkbook()
# if(grepl( "FDR >", web_call_BP(genes_sig_Up)))
up_bp <- no_web_call_BP(genes_sig_Up)

if(is.null(nrow(up_bp)) == T){
  print("No significant results for up BP")
}else{
up_bp_plot  <- ggplot(up_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO NoRedund (UP)", x = "enrichmentRatio", y = "Biological Process")

UpBp_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_NRED_biological_processes_GO.png")
# ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300)
# ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300, device = "png")
ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

file.exists(UpBp_path)

if("UP_BP_noRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, sheetName = "UP_BP_noRedund")
}
writeData(wb_GOredund, "UP_BP_noRedund", up_bp)
 
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_bp$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_bp$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)

  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_bp_NOREDUND_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_bp_NOREDUND_list")
}
  writeData(wb_GOredund, "UP_bp_NOREDUND_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

  
  print(up_bp_plot)
}

down_bp <- no_web_call_BP(genes_sig_down)

if(is.null(nrow(down_bp)) == T){
  print("No significant results for down BP")
}else{
down_bp_plot  <- ggplot(down_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO NoRedund (DOWN)", x = "enrichmentRatio", y = "Biological Process")

DownBp_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_NRED_biological_processes_GO.png")
# ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300)
ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
if("DWN_BP_noRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, sheetName = "DWN_BP_noRedund")
}
writeData(wb_GOredund, "DWN_BP_noRedund", down_bp)
 
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_bp$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_bp$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)

  
  
  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
if("DWN_BP_noRedund_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_BP_noRedund_list")
}
  writeData(wb_GOredund, "DWN_BP_noRedund_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

  print(down_bp_plot)
}

# up_bp_plot
# down_bp_plot
```

```{r}
#GO analysis for MF (UP and DOWN)

up_MF <- no_web_call_MF(genes_sig_Up)

if(is.null(nrow(up_MF)) == T){
  print("No significant results for up MF")
}else{
up_MF_plot  <- ggplot(up_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Molecular Function GO NoRedund (UP)", x = "enrichmentRatio", y = "Molecular Function")

UpMF_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_NRED_molecular_function_GO.png")
# ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("UP_MF_noRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, sheetName = "UP_MF_noRedund")
}
writeData(wb_GOredund, "UP_MF_noRedund", up_MF)
 
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_MF$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_MF$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_MF_noRedund_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_MF_noRedund_list")
}
  writeData(wb_GOredund, "UP_MF_noRedund_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

  print(up_MF_plot)
}

down_MF <- no_web_call_MF(genes_sig_down)

if(is.null(nrow(down_MF)) == T){
  print("No significant results for down MF")
}else{
down_MF_plot  <- ggplot(down_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Molecular Function GO NoRedund (DOWN)", x = "enrichmentRatio", y = "Molecular Function")

DownMF_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_NRED_molecular_function_GO.png")
# ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("DWN_MF_noRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, sheetName = "DWN_MF_noRedund")
}
writeData(wb_GOredund, "DWN_MF_noRedund", down_MF)
 
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_MF$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_MF$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)

  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_MF_noRedund_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_MF_noRedund_list")
}
  writeData(wb_GOredund, "DWN_MF_noRedund_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

  print(down_MF_plot)

}

# up_MF_plot
# down_MF_plot
```



#GO redundant
```{r}
#set up WebGestaltR function for BP
if(org == "human"){
web_call_BP <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process'){
#set if for organism type                                                                                                                     
  WebGestaltR(
    enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}
}else if(org == "mouse"){
  web_call_BP <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id",
                       path = 'geneontology_Molecular_Function'){
 #set if for organism type                                                                                                                     
  WebGestaltR(
    enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = sigAll, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}
}
```


```{r}
#GO analysis for BP (UP and DOWN)

# if(grepl( "FDR >", web_call_BP(genes_sig_Up)))
up_bp <- web_call_BP(genes_sig_Up)

if(is.null(nrow(up_bp)) == T){
  print("No significant results for up BP")
}else{
up_bp_plot  <- ggplot(up_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (UP)", x = "enrichmentRatio", y = "Biological Process")

UpBp_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_biological_processes_GO.png")
# ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300)

ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

 if("UP_GO_bpRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_bpRedund")
}
  writeData(wb_GOredund, "UP_GO_bpRedund", up_bp)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_bp$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_bp$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)

  
  
  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  if("UP_GO_bp_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_bp_list")
}
  writeData(wb_GOredund, "UP_GO_bp_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

  print(up_bp_plot)
}

down_bp <- web_call_BP(genes_sig_down)

if(is.null(nrow(down_bp)) == T){
  print("No significant results for down BP")
}else{
down_bp_plot  <- ggplot(down_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (DOWN)", x = "enrichmentRatio", y = "Biological Process")

DownBp_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_biological_processes_GO.png")
# ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300)

ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("DWN_GO_bpRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "DWN_GO_bpRedund")
}
  writeData(wb_GOredund, "DWN_GO_bpRedund", down_bp)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_bp$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_bp$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  

  # Apply symbol conversion across every column of the df
  GO_bp_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_bp_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_bp_list")
}
  writeData(wb_GOredund, "DWN_GO_bp_list", GO_bp_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  print(down_bp_plot)
}

# up_bp_plot
# down_bp_plot
```

```{r}
#GO analysis for MF (UP and DOWN)

up_MF <- web_call_MF(genes_sig_Up)

if(is.null(nrow(up_MF)) == T){
  print("No significant results for up MF")
}else{
up_MF_plot  <- ggplot(up_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (UP)", x = "enrichmentRatio", y = "Molecular Function")

UpMF_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_molecular_function_GO.png")
# ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("UP_GO_mfRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "UP_GO_mfRedund")
}
  writeData(wb_GOredund, "UP_GO_mfRedund", up_MF)

  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_MF$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_MF$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_MF_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_GO_MF_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_MF_list")
}
  writeData(wb_GOredund, "UP_GO_MF_list", GO_MF_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  print(up_MF_plot)
}

down_MF <- web_call_MF(genes_sig_down)

if(is.null(nrow(down_MF)) == T){
  print("No significant results for down MF")
}else{
down_MF_plot  <- ggplot(down_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (DOWN)", x = "enrichmentRatio", y = "Molecular Function")

DownMF_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_molecular_function_GO.png")
# ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("DWN_GO_mfRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "DWN_GO_mfRedund")
}
  writeData(wb_GOredund, "DWN_GO_mfRedund", down_MF)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_MF$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_MF$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_MF_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_MF_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_MF_list")
}
  writeData(wb_GOredund, "DWN_GO_MF_list", GO_MF_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  print(down_MF_plot)
}

# up_MF_plot
# down_MF_plot
```


```{r}
#GO analysis for KEGG (UP and DOWN)

up_KEGG <- no_web_call_KEGG(genes_sig_Up)

if(is.null(nrow(up_KEGG)) == T){
  print("No significant results")
}else{
up_KEGG_plot  <- ggplot(up_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "KEGG GO analysis (UP)", x = "enrichmentRatio", y = "KEGG")

UpKEGG_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_KEGG_GO.png")
# ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 12, height = 4, dpi = 300)
ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("UP_GO_KEGG" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "UP_GO_KEGG")
}
  writeData(wb_GOredund, "UP_GO_KEGG", up_KEGG)

library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_KEGG$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_KEGG$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_KEGG_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_GO_KEGG_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_KEGG_list")
}
  writeData(wb_GOredund, "UP_GO_KEGG_list", GO_KEGG_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

print(up_KEGG_plot)
}

down_KEGG <- no_web_call_KEGG(genes_sig_down)

if(is.null(nrow(down_KEGG)) == T){
  print("No significant results")
}else{
down_KEGG_plot  <- ggplot(down_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "KEGG GO analysis (DOWN)", x = "enrichmentRatio", y = "KEGG")

DownKEGG_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_KEGG_GO.png")
# ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 12, height = 4, dpi = 300)
ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("DWN_GO_KEGGRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "DWN_GO_KEGGRedund")
}
  writeData(wb_GOredund, "DWN_GO_KEGGRedund", down_KEGG)

library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_KEGG$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_KEGG$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_KEGG_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_KEGG_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_KEGG_list")
}
  writeData(wb_GOredund, "DWN_GO_KEGG_list", GO_KEGG_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

print(down_KEGG_plot)
}

```

```{r}
#GO analysis for Hallmark50 (UP and DOWN)
if(org == "human"){
up_Hall <- no_web_call_Hall(genes_sig_Up)

if(is.null(nrow(up_Hall)) == T){
  print("No significant results")
}else{
up_Hall_plot  <- ggplot(up_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

UpHall_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_Hallmarks_GO.png")
# ggsave(UpHall_path, plot = up_Hall_plot, width = 12, height = 4, dpi = 300)
ggsave(UpHall_path, plot = up_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("UP_GO_hallRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "UP_GO_hallRedund")
}
  writeData(wb_GOredund, "UP_GO_hallRedund", up_Hall)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_Hall$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_Hall$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_HALL_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_GO_HALL_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_HALL_list")
}
  writeData(wb_GOredund, "UP_GO_HALL_list", GO_HALL_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

print(up_Hall_plot)
}

down_Hall <- no_web_call_Hall(genes_sig_down)

if(is.null(nrow(down_Hall)) == T){
  print("No significant results")
}else{
down_Hall_plot  <- ggplot(down_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

DownHall_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_Hallmarks_GO.png")
# ggsave(DownHall_path, plot = down_Hall_plot, width = 12, height = 4, dpi = 300)
ggsave(DownHall_path, plot = down_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

if("DWN_GO_hallRedund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GOredund, "DWN_GO_hallRedund")
}
  writeData(wb_GOredund, "DWN_GO_hallRedund", down_Hall)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_Hall$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_Hall$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_HALL_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_HALL_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_HALL_list")
}
  writeData(wb_GOredund, "DWN_GO_HALL_list", GO_HALL_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)

print(down_Hall_plot)
}
}else{message("Hallmarks50 not available for mouse ORA")}
```


```{r}
#GO analysis for Panther (UP and DOWN)

up_Panth <- no_web_call_Panther(genes_sig_Up)

if(is.null(nrow(up_Panth)) == T){
  print("No significant results")
}else{
  up_Panth_plot  <- ggplot(up_Panth)+
    geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
    theme_classic() +
    labs(title = "Panther GO analysis (UP)", x = "enrichmentRatio", y = "Panther 50")
  
  if("UP_GO_PANTH_Redund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_PANTH_Redund")
}
  writeData(wb_GOredund, "UP_GO_PANTH_Redund", up_Panth)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_Panth$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_Panth$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_PANTH_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_GO_PANTH_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_PANTH_list")
}
  writeData(wb_GOredund, "UP_GO_PANTH_list", GO_PANTH_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  
  UpPanth_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_Panther_GO.png")
  # ggsave(UpPanth_path, plot = up_Panth_plot, width = 12, height = 4, dpi = 300)
  ggsave(UpPanth_path, plot = up_Panth_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
  
  print(up_Panth_plot)
}

down_Panther <- no_web_call_Panther(genes_sig_down)

if(is.null(nrow(down_Panther)) == T){
  print("No significant results")
}else{
  down_Panth_plot  <- ggplot(down_Panther)+
    geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
    theme_classic() +
    labs(title = "Panther GO analysis (UP)", x = "enrichmentRatio", y = "Panther 50")
  
  if("DWN_GO_PANTH_Redund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_PANTH_Redund")
}
  writeData(wb_GOredund, "DWN_GO_PANTH_Redund", down_Panther)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_Panther$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_Panther$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_PANTH_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_PANTH_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_PANTH_list")
}
  writeData(wb_GOredund, "DWN_GO_PANTH_list", GO_PANTH_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  
  DownPanth_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_Panther_GO.png")
  # ggsave(DownPanth_path, plot = down_Panth_plot, width = 12, height = 4, dpi = 300)
  ggsave(DownPanth_path, plot = down_Panth_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
  
  print(down_Panth_plot)
}



```

```{r}
up_React <- no_web_call_Reactome(genes_sig_Up)

if(is.null(nrow(up_React)) == T){
  print("No significant results")
}else{
  up_React_plot  <- ggplot(up_React)+
    geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
    theme_classic() +
    labs(title = "Reactome GO analysis (UP)", x = "enrichmentRatio", y = "Reactome")
  
  if("UP_GO_REACT_Redund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_REACT_Redund")
}
  writeData(wb_GOredund, "UP_GO_REACT_Redund", up_React)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(up_React$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- up_React$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_REACT_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("UP_GO_REACT_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "UP_GO_REACT_list")
}
  writeData(wb_GOredund, "UP_GO_REACT_list", GO_REACT_symbols)
  
  
  UpReact_path <- file.path(normalizePath(outputdir_SEQ_GO), "UP_Reactome_GO.png")
  # ggsave(UpReact_path, plot = up_React_plot, width = 12, height = 4, dpi = 300)
  ggsave(UpReact_path, plot = up_React_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
  
  print(up_React_plot)
}

down_React <- no_web_call_Reactome(genes_sig_down)

if(is.null(nrow(down_React)) == T){
  print("No significant results")
}else{
  down_React_plot  <- ggplot(down_React)+
    geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
    theme_classic() +
    labs(title = "Reactome GO analysis (DOWN)", x = "enrichmentRatio", y = "Reactome")
  
  if("DWN_GO_REACT_Redund" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_REACT_Redund")
}
  writeData(wb_GOredund, "DWN_GO_REACT_Redund", down_React)
  
  library(stringr)
  
  # Step 1: Split each semicolon-separated gene string into a vector
  gene_lists <- str_split(down_React$userId, ";")
  
  # Step 2: Name each list element by its GO term description
  names(gene_lists) <- down_Panther$description
  
  # Step 3: Pad shorter vectors with NA so all are equal length
  max_len <- max(lengths(gene_lists))
  gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))
  
  # Step 4: Coerce to data frame
  GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)
  
  library(org.Mm.eg.db)
  
  # Apply symbol conversion across every column of the df
  GO_REACT_symbols <- as.data.frame(
    lapply(GO_bp_df, function(col) {
      mapped <- mapIds(id_org,
                       keys = na.omit(col),
                       column = "SYMBOL",
                       keytype = "ENSEMBL",
                       multiVals = "first")
      # Re-insert NAs in correct positions
      result <- rep(NA, length(col))
      result[!is.na(col)] <- mapped
      return(result)
    }),
    check.names = FALSE
  )
  
  if("DWN_GO_REACT_list" %in% names(wb_GOredund)){
  print("updating current worksheet")
}else{
  addWorksheet(wb_GOredund, "DWN_GO_REACT_list")
}
  writeData(wb_GOredund, "DWN_GO_REACT_list", GO_REACT_symbols)
  # saveWorkbook(wb_GOredund, file = paste0(outputdir_SEQ, "/", cell_line, "_ORA_lists.xlsx"), overwrite = T)
  
  
  DownReact_path <- file.path(normalizePath(outputdir_SEQ_GO), "DWN_Reactome_GO.png")
  # ggsave(DownReact_path, plot = down_React_plot, width = 12, height = 4, dpi = 300)
  ggsave(DownReact_path, plot = down_React_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
  
  print(down_React_plot)
}



```


```{r}
saveWorkbook(wb_GOredund, file = file.path(outputdir_SEQ, paste0(cell_line, "_", perturbation,"_ORA_lists.xlsx")), overwrite = T)
```

#GSEA setup

```{r}
###############
#setup GSEA webgestaltR functions
##############

if(org == "human"){
#set up WebGestaltR function for BP
GSEA_call_BP <- function(x, cutoff = 0.05, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
  )
}

#WebGestaltR function for MF
GSEA_call_MF <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
GSEA_call_KEGG <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,


    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
GSEA_call_Hall <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
}
#WebGestaltR function for Panther
GSEA_call_Panth <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_Panther'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
} 

#WebGestaltR function for Reactome
GSEA_call_React <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_Reactome'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
} 

}else if (org == "mouse"){
 #set up WebGestaltR function for BP
GSEA_call_BP <- function(x, cutoff = 0.05, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "GSEA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
  )
}

#WebGestaltR function for MF
GSEA_call_MF <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
GSEA_call_KEGG <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,


    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Panther
GSEA_call_Panth <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_Panther'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
} 
#WebGestaltR function for Reactome
GSEA_call_React <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_Reactome'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
} 

}
```

#Signed pvalue GSEA
```{r}
library(DESeq2)
library(org.Hs.eg.db)
library(AnnotationDbi)

res_raw <- DESeq2::results(dds)

res_raw = lfcShrink(dds, contrast = c("Group","KD","Ctrl"), type = "ashr")

res_raw$ENSEMBL <- gsub("\\..*", "", rownames(res_raw))

res_raw <- as.data.frame(res_raw)



res_raw = res_raw[!is.na(res_raw$ENSEMBL) & !is.na(res_raw$log2FoldChange), ]


###############################
# #using stat from DESeq output
###############################
# gene_df <- data.frame(
#   gene = res_raw$ENSEMBL,
#   L2fc_score = res_raw$stat
# ) 
#################################
#signed pvalue
#################################
if(TRUE %in% is.na(res_raw$pvalue) == T){
    message("NA's found")
  }else{
    res_raw = res_raw %>%
      mutate(scores = sign(log2FoldChange) * abs(log10(pvalue)))
  }


gene_list <- res_raw$scores
names(gene_list) <- as.character(res_raw$ENSEMBL)
gene_list <- sort(gene_list, decreasing = TRUE)

gene_df <- data.frame(
  gene = res_raw$ENSEMBL,
  L2fc_score = res_raw$scores
)

gene_df = sort_by(gene_df, gene_df$L2fc_score, decreasing = T)

gene_df_name = gene_df 

gene_df_name$symbol = mapIds(id_org,
                             keys = gene_df_name$gene,
                             column = "SYMBOL",
                             keytype = "ENSEMBL",
                             multiVals = "first")

```


```{r}
GSEA_BP <- GSEA_call_BP(gene_df)

GSEA_MF <- GSEA_call_MF(gene_df)

GSEA_KEGG <- GSEA_call_KEGG(gene_df)
if(org == "human"){
GSEA_Hall <- GSEA_call_Hall(gene_df)
}else{
  message("no GSEA Hallmarks50 available for mouse")
}
GSEA_Panth = GSEA_call_Panth(gene_df)

GSEA_React = GSEA_call_React(gene_df)

GSEABP_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_BP.png")
GSEAMF_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_MF.png")
GSEAKEGG_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_KEGG.png")
GSEAHall_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_Hallmarks.png")
GSEAPanth_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_Panther.png")
GSEAReact_path <- file.path(normalizePath(outputdir_SEQ_GSEA_SignP), "GSEA_React.png")


wb_GSEA_signP = createWorkbook()

if("Gene_score_input" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "Gene_score_input")
}
writeData(wb_GSEA_signP, "Gene_score_input", gene_df_name)

```

```{r}
#GSEA BP output

# library(fgsea)
library(ggplot2)
if("BP_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "BP_GSEA")
}

if(is.null(nrow(GSEA_BP)) == T){
  print("No significant results for GSEA BP")
}else{

writeData(wb_GSEA_signP, "BP_GSEA", GSEA_BP)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_BP$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_BP$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)
if("BP_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "BP_geneList")
}
writeData(wb_GSEA_signP, "BP_geneList", GSEA_symbols)
  
GSEA_BP_plot <- ggplot(GSEA_BP, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Biological Processes",
    x = "Normalized Enrichment Score",
    y = "Biological Processes",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEABP_path, plot = GSEA_BP_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEABP_path, plot = GSEA_BP_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_BP_plot)

# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_BP.pdf")
# 
# 
#     
#   p = ggplot(GSEA_BP, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
#   
#   ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)
#     
#  

}
```

```{r}
#GSEA MF output
if("MF_GSEA" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "MF_GSEA")
}

if(is.null(nrow(GSEA_MF)) == T){
  print("No significant results for GSEA MF")
}else{
  
  writeData(wb_GSEA_signP, "MF_GSEA", GSEA_MF)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_MF$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_MF$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("MF_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "MF_geneList")
}
writeData(wb_GSEA_signP, "MF_geneList", GSEA_symbols)
  
GSEA_MF_plot <- ggplot(GSEA_MF, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Molecular Function",
    x = "Normalized Enrichment Score",
    y = "Molecular Function",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAMF_path, plot = GSEA_MF_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAMF_path, plot = GSEA_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_MF_plot)
# 
# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_MF.pdf")
# 
# p = ggplot(GSEA_MF_plot, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
# 
# ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)


}
```

```{r}
#GSEA KEGG output
if("KEGG_GSEA" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "KEGG_GSEA")
}
  
if(is.null(nrow(GSEA_KEGG)) == T){
  print("No significant results for GSEA KEGG")
}else{
  
  writeData(wb_GSEA_signP, "KEGG_GSEA", GSEA_KEGG)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_KEGG$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_KEGG$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("KEGG_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "KEGG_geneList")
}
writeData(wb_GSEA_signP, "KEGG_geneList", GSEA_symbols)
  
GSEA_KEGG_plot <- ggplot(GSEA_KEGG, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA KEGG",
    x = "Normalized Enrichment Score",
    y = "KEGG",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAKEGG_path, plot = GSEA_KEGG_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAKEGG_path, plot = GSEA_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_KEGG.pdf")
# 
# p = ggplot(GSEA_MF_plot, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
# 
# ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)

}
```

```{r}
#GSEA Hallmarks 50 output
if(org == "human"){
  if("Hall_GSEA" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "Hall_GSEA")
}
if(is.null(nrow(GSEA_Hall)) == T & org == "human"){
  print("No significant results for GSEA Hallmarks 50")
}else{
  
  writeData(wb_GSEA_signP, "Hall_GSEA", GSEA_Hall)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_Hall$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_Hall$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("Hall_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "Hall_geneList")
}
writeData(wb_GSEA_signP, "Hall_geneList", GSEA_symbols)
  
GSEA_Hall_plot <- ggplot(GSEA_Hall, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Hallmarks 50",
    x = "Normalized Enrichment Score",
    y = "Hallmarks",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAHall_path, plot = GSEA_Hall_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAHall_path, plot = GSEA_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_Hall_plot)
}

}else{ message("no Hallmarks50 available for mouse")}
# GSEA_BP_plot
# GSEA_MF_plot
# GSEA_KEGG_plot
# GSEA_Hall_plot
```


```{r}
#GSEA Panther
if("Panth_GSEA" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "Panth_GSEA")
}

if(is.null(nrow(GSEA_Panth)) == T){
  print("No significant results for GSEA Panther")
}else{
  
  writeData(wb_GSEA_signP, "Panth_GSEA", GSEA_Panth)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_Panth$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_Panth$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("Panth_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "Panth_geneList")
}
writeData(wb_GSEA_signP, "Panth_geneList", GSEA_symbols)
  
GSEA_Panth_plot <- ggplot(GSEA_Panth, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Panther",
    x = "Normalized Enrichment Score",
    y = "Panther",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAPanth_path, plot = GSEA_Panth_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAPanth_path, plot = GSEA_Panth_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_Panth_plot)
}

```

```{r}
#GSEA Reactome
if("React_GSEA" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "React_GSEA")
}

if(is.null(nrow(GSEA_React)) == T){
  print("No significant results for GSEA Reactome")
}else{
  
  writeData(wb_GSEA_signP, "React_GSEA", GSEA_React)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_React$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_React$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("React_geneList" %in% names(wb_GSEA_signP)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_signP, "React_geneList")
}
writeData(wb_GSEA_signP, "React_geneList", GSEA_symbols)
  
GSEA_React_plot <- ggplot(GSEA_React, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Reactome",
    x = "Normalized Enrichment Score",
    y = "Reactome",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAReact_path, plot = GSEA_React_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAReact_path, plot = GSEA_React_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_React_plot)
}

```

```{r}
#save signed pvalue GSEA workbook

saveWorkbook(wb_GSEA_signP, file = file.path(outputdir_SEQ, paste0(cell_line, "_", perturbation, "_SEQ_GSEA_signed_Pvalue.xlsx")), overwrite = T)

# rm(res_raw, gene_df)
```


#STAT GSEA
```{r}
library(DESeq2)
library(org.Hs.eg.db)
library(AnnotationDbi)

res_raw <- DESeq2::results(dds)

res_raw = lfcShrink(dds, contrast = c("Group","KD","Ctrl"), type = "ashr")

res_raw$ENSEMBL <- gsub("\\..*", "", rownames(res_raw))

res_raw <- as.data.frame(res_raw)


res_raw = res_raw[!is.na(res_raw$ENSEMBL) & !is.na(res_raw$log2FoldChange), ]

###############################
# #using stat from DESeq output
###############################
gene_df <- data.frame(
  gene = res_raw$ENSEMBL,
  L2fc_score = res_raw$stat
)

gene_list <- res_raw$L2fc_score
names(gene_list) <- as.character(res_raw$ENSEMBL)
gene_list <- sort(gene_list, decreasing = TRUE)


gene_df = sort_by(gene_df, gene_df$L2fc_score, decreasing = T)

gene_df_name = gene_df 

gene_df_name$symbol = mapIds(id_org,
                             keys = gene_df_name$gene,
                             column = "SYMBOL",
                             keytype = "ENSEMBL",
                             multiVals = "first")

```


```{r}
GSEA_BP <- GSEA_call_BP(gene_df)

GSEA_MF <- GSEA_call_MF(gene_df)

GSEA_KEGG <- GSEA_call_KEGG(gene_df)
if(org == "human"){
GSEA_Hall <- GSEA_call_Hall(gene_df)
}else{
  message("no GSEA Hallmarks50 available for mouse")
}
GSEA_Panth = GSEA_call_Panth(gene_df)

GSEA_React = GSEA_call_React(gene_df)

GSEABP_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_BP.png")
GSEAMF_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_MF.png")
GSEAKEGG_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_KEGG.png")
GSEAHall_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_Hallmarks.png")
GSEAPanth_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_Panther.png")
GSEAReact_path <- file.path(normalizePath(outputdir_SEQ_GSEA_STAT), "GSEA_React.png")


wb_GSEA_STAT = createWorkbook()

if("Gene_score_input" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "Gene_score_input")
}
writeData(wb_GSEA_STAT, "Gene_score_input", gene_df_name)

```

```{r}
#GSEA BP output

# library(fgsea)
library(ggplot2)
if("BP_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "BP_GSEA")
}

if(is.null(nrow(GSEA_BP)) == T){
  print("No significant results for GSEA BP")
}else{

writeData(wb_GSEA_STAT, "BP_GSEA", GSEA_BP)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_BP$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_BP$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)
if("BP_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "BP_geneList")
}
writeData(wb_GSEA_STAT, "BP_geneList", GSEA_symbols)
  
GSEA_BP_plot <- ggplot(GSEA_BP, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Biological Processes",
    x = "Normalized Enrichment Score",
    y = "Biological Processes",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEABP_path, plot = GSEA_BP_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEABP_path, plot = GSEA_BP_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_BP_plot)

# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_BP.pdf")
# 
# 
#     
#   p = ggplot(GSEA_BP, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
#   
#   ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)
#     
#  

}
```

```{r}
#GSEA MF output
if("MF_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "MF_GSEA")
}

if(is.null(nrow(GSEA_MF)) == T){
  print("No significant results for GSEA MF")
}else{
  
  writeData(wb_GSEA_STAT, "MF_GSEA", GSEA_MF)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_MF$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_MF$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("MF_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "MF_geneList")
}
writeData(wb_GSEA_STAT, "MF_geneList", GSEA_symbols)
  
GSEA_MF_plot <- ggplot(GSEA_MF, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Molecular Function",
    x = "Normalized Enrichment Score",
    y = "Molecular Function",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAMF_path, plot = GSEA_MF_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAMF_path, plot = GSEA_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_MF_plot)
# 
# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_MF.pdf")
# 
# p = ggplot(GSEA_MF_plot, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
# 
# ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)


}
```

```{r}
#GSEA KEGG output
if("KEGG_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "KEGG_GSEA")
}
  
if(is.null(nrow(GSEA_KEGG)) == T){
  print("No significant results for GSEA KEGG")
}else{
  
  writeData(wb_GSEA_STAT, "KEGG_GSEA", GSEA_KEGG)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_KEGG$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_KEGG$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("KEGG_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "KEGG_geneList")
}
writeData(wb_GSEA_STAT, "KEGG_geneList", GSEA_symbols)
  
GSEA_KEGG_plot <- ggplot(GSEA_KEGG, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA KEGG",
    x = "Normalized Enrichment Score",
    y = "KEGG",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAKEGG_path, plot = GSEA_KEGG_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAKEGG_path, plot = GSEA_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

# output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_KEGG.pdf")
# 
# p = ggplot(GSEA_MF_plot, aes(
#     x = normalizedEnrichmentScore,
#     y = reorder(description, normalizedEnrichmentScore),
#     size = size,
#     color = -log10(FDR)
# ))
# 
# ggsave(output_seqGSEA, plot = p, width = 12, height = 4, dpi = 300, 
#        device = grDevices::png)

}
```

```{r}
#GSEA Hallmarks 50 output
if(org == "human"){
  if("Hall_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "Hall_GSEA")
}
if(is.null(nrow(GSEA_Hall)) == T & org == "human"){
  print("No significant results for GSEA Hallmarks 50")
}else{
  
  writeData(wb_GSEA_STAT, "Hall_GSEA", GSEA_Hall)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_Hall$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_Hall$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("Hall_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "Hall_geneList")
}
writeData(wb_GSEA_STAT, "Hall_geneList", GSEA_symbols)
  
GSEA_Hall_plot <- ggplot(GSEA_Hall, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Hallmarks 50",
    x = "Normalized Enrichment Score",
    y = "Hallmarks",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAHall_path, plot = GSEA_Hall_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAHall_path, plot = GSEA_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_Hall_plot)
}

}else{ message("no Hallmarks50 available for mouse")}
# GSEA_BP_plot
# GSEA_MF_plot
# GSEA_KEGG_plot
# GSEA_Hall_plot
```


```{r}
#GSEA Panther
if("Panth_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "Panth_GSEA")
}

if(is.null(nrow(GSEA_Panth)) == T){
  print("No significant results for GSEA Panther")
}else{
  
  writeData(wb_GSEA_STAT, "Panth_GSEA", GSEA_Panth)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_Panth$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_Panth$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("Panth_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "Panth_geneList")
}
writeData(wb_GSEA_STAT, "Panth_geneList", GSEA_symbols)
  
GSEA_Panth_plot <- ggplot(GSEA_Panth, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Panther",
    x = "Normalized Enrichment Score",
    y = "Panther",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAPanth_path, plot = GSEA_Panth_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAPanth_path, plot = GSEA_Panth_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_Panth_plot)
}

```

```{r}
#GSEA Reactome
if("React_GSEA" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "React_GSEA")
}

if(is.null(nrow(GSEA_React)) == T){
  print("No significant results for GSEA Reactome")
}else{
  
  writeData(wb_GSEA_STAT, "React_GSEA", GSEA_React)

# Step 1: Split each semicolon-separated gene string into a vector
gene_lists <- str_split(GSEA_React$userId, ";")

# Step 2: Name each list element by its GO term description
names(gene_lists) <- GSEA_React$description

# Step 3: Pad shorter vectors with NA so all are equal length
max_len <- max(lengths(gene_lists))
gene_lists_padded <- lapply(gene_lists, function(x) c(x, rep(NA, max_len - length(x))))

# Step 4: Coerce to data frame
GO_bp_df <- as.data.frame(gene_lists_padded, check.names = FALSE)


GSEA_symbols <- as.data.frame(
  lapply(GO_bp_df, function(col) {
    mapped <- mapIds(id_org,
                     keys = na.omit(col),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")
    # Re-insert NAs in correct positions
    result <- rep(NA, length(col))
    result[!is.na(col)] <- mapped
    return(result)
  }),
  check.names = FALSE
)

if("React_geneList" %in% names(wb_GSEA_STAT)){
  print("updating current worksheet")
}else{
addWorksheet(wb_GSEA_STAT, "React_geneList")
}
writeData(wb_GSEA_STAT, "React_geneList", GSEA_symbols)
  
GSEA_React_plot <- ggplot(GSEA_React, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Reactome",
    x = "Normalized Enrichment Score",
    y = "Reactome",
    size = "Gene count",
    color = "FDR"
  )

# ggsave(GSEAReact_path, plot = GSEA_React_plot, width = 12, height = 6, dpi = 300)
ggsave(GSEAReact_path, plot = GSEA_React_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

print(GSEA_React_plot)
}

```

```{r}
#save STAT ranking GSEA workbook

saveWorkbook(wb_GSEA_STAT, file = file.path(outputdir_SEQ, paste0(cell_line, "_", perturbation, "_SEQ_GSEA_STAT.xlsx")), overwrite = T)

```

#ggsashimi DESEq2 top 100 events 
```{r, warning=FALSE, message=FALSE}
#generate ggsashimi plots for top 100 events abs(dPSI) > 0.01 and deposit to rMATS folder
library(ggplot2)
library(data.table)
library(gridExtra)

#generate table for ggsashimi DESEq run

SEQ_top100_temp <- res_df_allfilt[order(-res_df_allfilt$log2FoldChange),]

SEQ_top100 <- bind_rows(
  slice_head(SEQ_top100_temp, n = 50),
  slice_tail(SEQ_top100_temp, n = 50)
)


res_ggsashimi_genes <- SEQ_top100$ENSEMBL


gtf_SEQ = gtf[which(gtf$feature == "gene"),]

gene_info <- lapply(res_ggsashimi_genes, function(gene){

  row <- gtf_SEQ[
    grepl(paste0("gene_id ", gene), gtf_SEQ$attribute) &
    gtf_SEQ$feature == "gene",
  ][1,]

  list(
    region = paste0(row$chr, ":", row$start, "-", row$end),
    gene_name = sub('.*gene_name ([^;]+);.*', '\\1', row$attribute)
  )
})

DEseq_regions <- sapply(gene_info, '[[', "region")
Seq_genes <- sapply(gene_info, '[[', "gene_name")


colors_key <- c(rep("#1f77b4",length(ctrl_bams2)), rep("#ff7f0e", length(KD_bams2)))
# reps <- 1:Replicates
# rep_nums <- rep(reps, each = 1)
# Ungrp <- unfactor(Group) %>% paste0(rep_nums)

ctrl_reps = 1:length(ctrl_bams2)
KD_reps = 1:length(KD_bams2)

ctrl_nums <- rep(ctrl_reps, each = 1)
KD_reps <- rep(KD_reps, each = 1)

Ungrp_ctrl <- c("ctrl") %>% paste0(ctrl_reps)
Ungrp_KD <- c("KD") %>% paste0(KD_reps)

Ungrp = c(Ungrp_ctrl, Ungrp_KD)

bams_docked <- sub(".*/", "/bams/", gg.bams)
bams.tsv <-  data.frame(group = Ungrp, bams = bams_docked, colorsK = colors_key )
#fixed /bams/... issue with bams.tsv output
write_tsv(bams.tsv, file.path(outputdir_base, "gg_bams.tsv"))



ggsashimi_path <- path.expand("~/ggsashimi/ggsashimi.py")
bam_tsv <- normalizePath(file.path(outputdir_base, "gg_bams.tsv"))

gtf_splitr <- function(x){
  parts <- str_split(x, pattern = "/")
  
  paste(parts[[1]][-length(parts[[1]])], collapse = '/')
}
gtf1_path <- gtf_splitr(gtf.path)
gtf_name <- str_split_i(gtf.path, pattern = "/", -1)
bam_tsv_name <- str_split_i(bam_tsv, pattern = "/", -1)



cmd <- "docker"

for(i in seq_along(DEseq_regions)){
  region <- DEseq_regions[i]
  gene_name <- Seq_genes[i]
  
  safe_region <- gsub("[:\\-]", "_", region)
  safe_name <- paste0(i, "_",safe_region, "_", gene_name)
  print(safe_name)
  docker_args <- list(
  c("run", "--rm"),
  c("-v", paste0(normalizePath(outputdir_base), "/:/tsv")),
  c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")),
  c("-v", paste0(normalizePath(BAM_files), "/:/bams")),
  c("-v", paste0(normalizePath(outputdir_SEQ_allsashimi), "/:/out")),
  c("-w", "/out"),
  "guigolab/ggsashimi",
  c("-b", paste0("/tsv/", bam_tsv_name)),
  c("-c", region),
  c("-g", paste0("/gtf/", gtf_name)),
  c("-o", safe_name),
  c("-C", "3"),
  c("-M", "5000"),
  "--fix-y-scale"
)


  
  tryCatch( 
    system2(cmd, args = unlist(docker_args),stdout = TRUE, stderr = TRUE),
    error = function(e){
      message("Failed region: ", region)
    }
  )
}
  

```


```{r}
DESEQ_sashimi <- list.files(outputdir_SEQ_allsashimi, pattern = "\\.pdf$", full.names = TRUE)
DESEQ_sashimi <- DESEQ_sashimi[mixedorder(DESEQ_sashimi)]

#order of gene names for labeled merged pdf
genes <- sapply(DESEQ_sashimi, function(name_path){
  sub('.*\\d_([^_]+)\\..*', "\\1", name_path)
}) %>% as.vector()

output_file <- normalizePath(
  file.path(outputdir_SEQ, "temp_100Events.pdf"),
  mustWork = FALSE
)


library(pdftools)
dir.create(dirname(output_file), recursive = TRUE, showWarnings = FALSE)
pdf_combine(DESEQ_sashimi, output = output_file)

library(pdftools)
library(grid)
library(grDevices)

merged_pdf <- output_file

n_pages <- pdf_info(merged_pdf)$pages

```

#label and merge Seq sashimi plots for final pdf
```{r, results='hide'}
library(pdftools)
library(png)
library(grid)

merged_pdf <- output_file
n_pages <- pdf_info(merged_pdf)$pages


tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)

annotated_pdfs <- character(n_pages)

for(i in seq_along(tmp_pngs)){

  img <- png::readPNG(tmp_pngs[i])

  out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
  annotated_pdfs[i] <- out_pdf

  pdf(out_pdf, width = 8, height = 6)

  grid.raster(img,
              x = 0.5, y = 0.5,
              width = unit(1, "npc"),
              height = unit(1, "npc"))

  grid.text(genes[i],
            x = 0.5, y = 0.97,
            gp = gpar(fontsize = 16, fontface = "bold"))

  dev.off()
}


final_pdf <- file.path(outputdir_SEQ, "top_100_events_labeled.pdf")

pdf_combine(annotated_pdfs, output = final_pdf)
```



#Run rMATS in terminal
```{r}
if(rMATS_runner == T){
rMATS_path <- path.expand("/home/bree/miniforge3/envs/rmats_env/bin/rmats.py")
# system2(path.expand("/home/bree/miniforge3/envs/rmats_env/bin/python"), args = c(
#   rMATS_path,
#   "--b1", ctrl_bams.path,
#   "--b2", KD_bams.path,
#   "--gtf", gtf.path,
#   "-t", "paired",
#   "--readLength", 150,
#   "--nthread", 12,
#   "--od", outputdir_MATSfiles,
#   "--tmp", outputdir_MATSfiles,
#   "--individual-counts", 
#   "--novelSS",
#   "--variable-read-length"
#   )
#   
# )
# }


  

cmd <- "docker"

#   docker_args_MATS <- list(
#   c("run", "--rm"), 
#   c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")), #good 
#   c("-v", paste0(normalizePath(BAM_files), "/:/bams")), #set to base rMATS folder path
#   c("-v", paste0(normalizePath(outputdir_MATS), "/:/out")), #set to rMATS_output folder path
#   c("-w", "/out"),
#   "xinglab/rmats",
#   c("python", "/rmats/rmats.py"),
#   c("--b1", paste0("/out/", ctrl_bams_name)), #set ctrl_bams_name
#   c("--b2", paste0("/out/", KD_bams_name)),  #set KD_bams_name
#   c("--gtf", paste0("/gtf/", gtf_name)),
#   c("-t", "paired"),
#   c("--readLength", 150),
#   c("--nthread", 34),
#   c("--od", paste0("/out/", "rMATS_output")),
#   c("--tmp", paste0("/out/", "rMATS_output")),
#   "--individual-counts", 
#   "--novelSS",
#   "--variable-read-length"
# 
# )

docker_args_MATS <- c(
  "run", "--rm",
  "-v", paste0(normalizePath(gtf1_path, mustWork = FALSE), ":/gtf"),
  "-v", paste0(normalizePath(BAM_files, mustWork = FALSE), ":/bams"),
  "-v", paste0(normalizePath(outputdir_MATS, mustWork = FALSE), ":/out"),
  "-w", "/out",
  "xinglab/rmats",
  "python", "/rmats/rmats.py",
  "--b1", paste0("/out/", ctrl_bams_name),
  "--b2", paste0("/out/", KD_bams_name),
  "--gtf", paste0("/gtf/", gtf_name),
  "-t", "paired",
  "--readLength", "150",
  "--nthread", "34",
  "--od", paste0("/out/", "rMATS_output"),
  "--tmp", paste0("/out/", "rMATS_output"),
  "--individual-counts",
  "--novelSS",
  "--variable-read-length"
)


 cat(paste(c(cmd, unlist(docker_args_MATS)), collapse = " "), "\n") 
  tryCatch({ 
    system2(cmd, args = unlist(docker_args_MATS),stdout = TRUE, stderr = TRUE)
},
    error = function(e){
      message("Failed region: ", region)
      message("Error message: ", e$message)
    }
  )
}
  
```


#Start excel for SE rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_M = paste0(cell_line, "_", perturbation, "_SE","_rMATS.xlsx")

outfile_M = paste(outputdir_MATS, fileName_M, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb2 <- createWorkbook()

if("README" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "README")
}
writeData(wb2, "README", README_df_M)
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)
```

#SE JCEC sheets
```{r}
#Raw rMATS output

#SE JCEC
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


JCEC.path <- list.files(
    MATS.path,
    pattern = "*SE.*MATS.*JCEC.txt$",
    full.names = TRUE
)

JCEC.path <- normalizePath(JCEC.path)
}else{
  JCEC.path <- list.files(
    MATS.path,
    pattern = "*SE.*MATS.*JCEC.txt$",
    full.names = TRUE
)

JCEC.path <- normalizePath(JCEC.path)
}

  
SE_MATS_JCEC <- tryCatch({
  read_delim(JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

#read in SE NovelSS
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.SE.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}else{
  NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.SE.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}

SE_NovelSS <- tryCatch({
  read_delim(NovelSS.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

SE_NovelSS <- SE_NovelSS %>%
  mutate(SpliceID = paste0(chr, exonStart_0base, exonEnd, upstreamES, upstreamEE, downstreamES, downstreamEE))

#read in SE novelJunction
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.SE.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}else{
  NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.SE.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}

SE_NovelJunc <- tryCatch({
  read_delim(NovelJunc.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

SE_NovelJunc <- SE_NovelJunc %>%
  mutate(SpliceID = paste0(chr, exonStart_0base, exonEnd, upstreamES, upstreamEE, downstreamES, downstreamEE))
  
SE_MATS_JCEC <- tryCatch({
  read_delim(JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

SE_MATS_JCEC <- SE_MATS_JCEC %>%
  mutate(SpliceID = paste0(chr, exonStart_0base, exonEnd, upstreamES, upstreamEE, downstreamES, downstreamEE))

#check if event is NovelSS
SE_MATS_JCEC <- SE_MATS_JCEC %>%
  mutate(NovelSS = ifelse(SE_MATS_JCEC$SpliceID %in% SE_NovelSS$SpliceID, "NovelSS", ""), .after = GeneID)
#check if event is NovelJunction
SE_MATS_JCEC <- SE_MATS_JCEC %>%
  mutate(NovelJunction = ifelse(SE_MATS_JCEC$SpliceID %in% SE_NovelJunc$SpliceID, "NovelJunction", ""), .after = NovelSS)

SE_MATS_JCEC <- SE_MATS_JCEC[order(SE_MATS_JCEC$FDR),]

if("SE_raw_rMATS_JCEC" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_raw_rMATS_JCEC")
}
writeData(wb2, "SE_raw_rMATS_JCEC",SE_MATS_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

SE_MATS_JCEC$GeneID <- gsub("\\..*", "", SE_MATS_JCEC$GeneID)

# MATS_geneList <- SE_MATS_JCEC$geneSymbol

SE_MATS_JCEC <- cbind(
  SE_MATS_JCEC,
  with_TPM[match(SE_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
SE_MATS_JCEC <- SE_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
SE_MATS_JCEC <- SE_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = upstream_to_downstream_count)
SE_MATS_JCEC<- SE_MATS_JCEC %>%
  mutate(query = GeneID)
SE_MATS_JCEC = SE_MATS_JCEC %>%
  mutate(IGV_coord = paste0(chr, ":", exonStart_0base, "-", exonEnd))

SE_MATS_JCEC <- SE_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("SE_rMATS_w_TPM" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_rMATS_w_TPM")
}
writeData(wb2, "SE_rMATS_w_TPM",SE_MATS_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

JC_cutoff <- 20

SJC_ctrl <- as.vector(str_split(SE_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
SJC_KD <- as.vector(str_split(SE_MATS_JCEC$SJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})

IJC_ctrl <- as.vector(str_split(SE_MATS_JCEC$IJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
IJC_KD <- as.vector(str_split(SE_MATS_JCEC$IJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})


SE_COUNT_JCEC <- SE_MATS_JCEC[which((SJC_ctrl >= JC_cutoff | SJC_KD >= JC_cutoff) & (IJC_ctrl >= JC_cutoff | IJC_KD >= JC_cutoff)), ]


if("SE_JC_cutoff_20" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_JC_cutoff_20")
}
writeData(wb2, "SE_JC_cutoff_20", SE_COUNT_JCEC)
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

#filter for padj (FDR) < 0.05
SE_PADJ_JCEC <- SE_COUNT_JCEC %>%
  filter(FDR < 0.05)

if("SE_padj_0.05" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_padj_0.05")
}
writeData(wb2, "SE_padj_0.05",SE_PADJ_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
SE_PSI_JCEC <- SE_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))
gene_info_MATS <- getGenes(geneids = SE_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

SE_PSI_JCEC <- SE_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

if("SE_dPSI_0.1" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_dPSI_0.1")
}
writeData(wb2, "SE_dPSI_0.1",SE_PSI_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

#separate JCEC file into canon and novel event sheets

SE_PSI_annotated <- SE_PSI_JCEC[which((SE_PSI_JCEC$NovelSS) %in% "" & (SE_PSI_JCEC$NovelJunction) %in% ""),]

if("SE_filt_annotated" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_filt_annotated")
}
writeData(wb2, "SE_filt_annotated", SE_PSI_annotated)
saveWorkbook(wb2, file = outfile_M, overwrite = T)

SE_PSI_novel <- SE_PSI_JCEC[which(!((SE_PSI_JCEC$NovelSS) %in% "") | !((SE_PSI_JCEC$NovelJunction) %in% "")),]

if("SE_filtered_novel" %in% names(wb2)){
  print("updating current worksheet")
}else{
addWorksheet(wb2, "SE_filtered_novel")
}
writeData(wb2, "SE_filtered_novel", SE_PSI_novel)
saveWorkbook(wb2, file = outfile_M, overwrite = T)



```

#Start MXE excel for rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_MXE = paste0(cell_line, "_", perturbation, "_MXE","_rMATS.xlsx")

outfile_MXE = paste(outputdir_MATS, fileName_MXE, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb3 <- createWorkbook()

if("README" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "README")
}
writeData(wb3, "README", README_df_M)
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)
```

#MXE JCEC sheets
```{r}
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


MXE_JCEC.path <- list.files(
    MATS.path,
    pattern = "*MXE.*MATS.*JCEC.txt$",
    full.names = TRUE
)

MXE_JCEC.path <- normalizePath(MXE_JCEC.path)
}else{
  MXE_JCEC.path <- list.files(
    MATS.path,
    pattern = "*MXE.*MATS.*JCEC.txt$",
    full.names = TRUE
)

MXE_JCEC.path <- normalizePath(MXE_JCEC.path)
}

  
MXE_MATS_JCEC <- tryCatch({
  read_delim(MXE_JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

MXE_MATS_JCEC <- MXE_MATS_JCEC %>%
  mutate(SpliceID = paste0(chr, `1stExonStart_0base`,`1stExonEnd`,`2ndExonStart_0base`,`2ndExonEnd`,upstreamES,upstreamEE,downstreamES,downstreamEE))

#read in SE NovelSS
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.MXE.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}else{
  NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.MXE.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}

MXE_NovelSS <- tryCatch({
  read_delim(NovelSS.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

MXE_NovelSS <- MXE_NovelSS %>%
  mutate(SpliceID = paste0(chr, `1stExonStart_0base`,`1stExonEnd`,`2ndExonStart_0base`,`2ndExonEnd`,upstreamES,upstreamEE,downstreamES,downstreamEE))

#read in SE novelJunction
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.MXE.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}else{
  NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.MXE.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}

MXE_NovelJunc <- tryCatch({
  read_delim(NovelJunc.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

MXE_NovelJunc <- MXE_NovelJunc %>%
  mutate(SpliceID = paste0(chr, `1stExonStart_0base`,`1stExonEnd`,`2ndExonStart_0base`,`2ndExonEnd`,upstreamES,upstreamEE,downstreamES,downstreamEE))


#check if event is NovelSS
MXE_MATS_JCEC <- MXE_MATS_JCEC %>%
  mutate(NovelSS = ifelse(MXE_MATS_JCEC$SpliceID %in% MXE_NovelSS$SpliceID, "NovelSS", ""), .after = GeneID)
#check if event is NovelJunction
MXE_MATS_JCEC <- MXE_MATS_JCEC %>%
  mutate(NovelJunction = ifelse(MXE_MATS_JCEC$SpliceID %in% MXE_NovelJunc$SpliceID, "NovelJunction", ""), .after = NovelSS)
MXE_MATS_JCEC <- MXE_MATS_JCEC[order(MXE_MATS_JCEC$FDR),]



if("MXE_raw_rMATS_JCEC" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_raw_rMATS_JCEC")
}
writeData(wb3, "MXE_raw_rMATS_JCEC",MXE_MATS_JCEC )
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)

MXE_MATS_JCEC$GeneID <- gsub("\\..*", "", MXE_MATS_JCEC$GeneID)


# MATS_geneList <- MXE_MATS_JCEC$geneSymbol

MXE_MATS_JCEC <- cbind(
  MXE_MATS_JCEC,
  with_TPM[match(MXE_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
MXE_MATS_JCEC <- MXE_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
MXE_MATS_JCEC <- MXE_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = second_count)
MXE_MATS_JCEC<- MXE_MATS_JCEC %>%
  mutate(query = GeneID)
MXE_MATS_JCEC = MXE_MATS_JCEC %>%
  mutate(IGV_coords = paste0(chr, ":", `1stExonStart_0base`, "-", `2ndExonEnd`))

MXE_MATS_JCEC <- MXE_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("MXE_rMATS_w_TPM" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_rMATS_w_TPM")
}
writeData(wb3, "MXE_rMATS_w_TPM",MXE_MATS_JCEC )
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)

JC_cutoff <- 20

SJC_ctrl <- as.vector(str_split(MXE_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
SJC_KD <- as.vector(str_split(MXE_MATS_JCEC$SJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})

IJC_ctrl <- as.vector(str_split(MXE_MATS_JCEC$IJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
IJC_KD <- as.vector(str_split(MXE_MATS_JCEC$IJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})


MXE_COUNT_JCEC <- MXE_MATS_JCEC[which((SJC_ctrl >= JC_cutoff | SJC_KD >= JC_cutoff) & (IJC_ctrl >= JC_cutoff | IJC_KD >= JC_cutoff)), ]

if("MXE_JC_cutoff_20" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_JC_cutoff_20")
}
writeData(wb3, "MXE_JC_cutoff_20", MXE_COUNT_JCEC)
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)

#filter for padj (FDR) < 0.05
MXE_PADJ_JCEC <- MXE_COUNT_JCEC %>%
  filter(FDR < 0.05)

if("MXE_padj_0.05" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_padj_0.05")
}
writeData(wb3, "MXE_padj_0.05",MXE_PADJ_JCEC )
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
MXE_PSI_JCEC <- MXE_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))

gene_info_MATS <- getGenes(geneids = MXE_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

MXE_PSI_JCEC <- MXE_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

if("MXE_dPSI_0.1" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_dPSI_0.1")
}
writeData(wb3, "MXE_dPSI_0.1",MXE_PSI_JCEC )
saveWorkbook(wb3, file = outfile_MXE, overwrite = TRUE)

#separate JCEC file into canon and novel event sheets

MXE_PSI_annotated <- MXE_PSI_JCEC[which((MXE_PSI_JCEC$NovelSS) %in% "" & (MXE_PSI_JCEC$NovelJunction) %in% ""),]

if("MXE_filt_annotated" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_filt_annotated")
}
writeData(wb3, "MXE_filt_annotated", MXE_PSI_annotated)
saveWorkbook(wb3, file = outfile_MXE, overwrite = T)

MXE_PSI_novel <- MXE_PSI_JCEC[which(!((MXE_PSI_JCEC$NovelSS) %in% "") | !((MXE_PSI_JCEC$NovelJunction) %in% "")),]

if("MXE_filtered_novel" %in% names(wb3)){
  print("updating current worksheet")
}else{
addWorksheet(wb3, "MXE_filtered_novel")
}
writeData(wb3, "MXE_filtered_novel", MXE_PSI_novel)
saveWorkbook(wb3, file = outfile_MXE, overwrite = T)

```

#Start RI excel for rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_RI = paste0(cell_line, "_", perturbation, "_RI","_rMATS.xlsx")

outfile_RI = paste(outputdir_MATS, fileName_RI, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb4 <- createWorkbook()

if("README" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "README")
}
writeData(wb4, "README", README_df_M)
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)
```

#RI JCEC sheets
```{r}
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


RI_JCEC.path <- list.files(
    MATS.path,
    pattern = "*RI.*MATS.*JCEC.txt$",
    full.names = TRUE
)

RI_JCEC.path <- normalizePath(RI_JCEC.path)
}else{
  RI_JCEC.path <- list.files(
    MATS.path,
    pattern = "*RI.*MATS.*JCEC.txt$",
    full.names = TRUE
)

RI_JCEC.path <- normalizePath(RI_JCEC.path)
}

  
RI_MATS_JCEC <- tryCatch({
  read_delim(RI_JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

RI_MATS_JCEC <- RI_MATS_JCEC %>%
  mutate(SpliceID = paste0(chr,riExonStart_0base,riExonEnd,upstreamES,upstreamEE,downstreamES,downstreamEE))

#read in RI NovelSS
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.RI.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}else{
  NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.RI.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}

RI_NovelSS <- tryCatch({
  read_delim(NovelSS.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

RI_NovelSS <- RI_NovelSS %>%
  mutate(SpliceID = paste0(chr,riExonStart_0base,riExonEnd,upstreamES,upstreamEE,downstreamES,downstreamEE))

#read in RI novelJunction
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.RI.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}else{
  NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.RI.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}

RI_NovelJunc <- tryCatch({
  read_delim(NovelJunc.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

RI_NovelJunc <- RI_NovelJunc %>%
  mutate(SpliceID = paste0(chr,riExonStart_0base,riExonEnd,upstreamES,upstreamEE,downstreamES,downstreamEE))


#check if event is NovelSS
RI_MATS_JCEC <- RI_MATS_JCEC %>%
  mutate(NovelSS = ifelse(RI_MATS_JCEC$SpliceID %in% RI_NovelSS$SpliceID, "NovelSS", ""), .after = GeneID)
#check if event is NovelJunction
RI_MATS_JCEC <- RI_MATS_JCEC %>%
  mutate(NovelJunction = ifelse(RI_MATS_JCEC$SpliceID %in% RI_NovelJunc$SpliceID, "NovelJunction", ""), .after = NovelSS)


RI_MATS_JCEC <- RI_MATS_JCEC[order(RI_MATS_JCEC$FDR),]


if("RI_raw_rMATS_JCEC" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_raw_rMATS_JCEC")
}
writeData(wb4, "RI_raw_rMATS_JCEC",RI_MATS_JCEC )
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)

RI_MATS_JCEC$GeneID <- gsub("\\..*", "", RI_MATS_JCEC$GeneID)


# MATS_geneList <- RI_MATS_JCEC$geneSymbol

RI_MATS_JCEC <- cbind(
  RI_MATS_JCEC,
  with_TPM[match(RI_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
RI_MATS_JCEC <- RI_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
RI_MATS_JCEC <- RI_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = upstream_to_downstream_count)
RI_MATS_JCEC<- RI_MATS_JCEC %>%
  mutate(query = GeneID)
RI_MATS_JCEC = RI_MATS_JCEC %>%
  mutate(IGV_coords = paste0(chr, ":", riExonStart_0base, "-", riExonEnd))

RI_MATS_JCEC <- RI_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("RI_rMATS_w_TPM" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_rMATS_w_TPM")
}
writeData(wb4, "RI_rMATS_w_TPM",RI_MATS_JCEC )
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)

JC_cutoff <- 20

SJC_ctrl <- as.vector(str_split(RI_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
SJC_KD <- as.vector(str_split(RI_MATS_JCEC$SJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})

IJC_ctrl <- as.vector(str_split(RI_MATS_JCEC$IJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
IJC_KD <- as.vector(str_split(RI_MATS_JCEC$IJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})


RI_COUNT_JCEC <- RI_MATS_JCEC[which((SJC_ctrl >= JC_cutoff | SJC_KD >= JC_cutoff) & (IJC_ctrl >= JC_cutoff | IJC_KD >= JC_cutoff)), ]

if("RI_JC_cutoff_20" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_JC_cutoff_20")
}
writeData(wb4, "RI_JC_cutoff_20", RI_COUNT_JCEC)
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)

#filter for padj (FDR) < 0.05
RI_PADJ_JCEC <- RI_COUNT_JCEC %>%
  filter(FDR < 0.05)

if("RI_padj_0.05" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_padj_0.05")
}
writeData(wb4, "RI_padj_0.05",RI_PADJ_JCEC )
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
RI_PSI_JCEC <- RI_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))

gene_info_MATS <- getGenes(geneids = RI_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

RI_PSI_JCEC <- RI_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

if("RI_dPSI_0.1" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_dPSI_0.1")
}
writeData(wb4, "RI_dPSI_0.1",RI_PSI_JCEC )
saveWorkbook(wb4, file = outfile_RI, overwrite = TRUE)


RI_PSI_annotated <- RI_PSI_JCEC[which((RI_PSI_JCEC$NovelSS) %in% "" & (RI_PSI_JCEC$NovelJunction) %in% ""),]

if("RI_filt_annotated" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_filt_annotated")
}
writeData(wb4, "RI_filt_annotated", RI_PSI_annotated)
saveWorkbook(wb4, file = outfile_RI, overwrite = T)

RI_PSI_novel <- RI_PSI_JCEC[which(!((RI_PSI_JCEC$NovelSS) %in% "") | !((RI_PSI_JCEC$NovelJunction) %in% "")),]

if("RI_filtered_novel" %in% names(wb4)){
  print("updating current worksheet")
}else{
addWorksheet(wb4, "RI_filtered_novel")
}
writeData(wb4, "RI_filtered_novel", RI_PSI_novel)
saveWorkbook(wb4, file = outfile_RI, overwrite = T)

```

#Start A3SS excel for rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_A3SS = paste0(cell_line, "_", perturbation, "_A3SS","_rMATS.xlsx")

outfile_A3SS = paste(outputdir_MATS, fileName_A3SS, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb5 <- createWorkbook()

if("README" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "README")
}
writeData(wb5, "README", README_df_M)
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)
```

#A3SS JCEC sheets
```{r}
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


A3SS_JCEC.path <- list.files(
    MATS.path,
    pattern = "*A3SS.*MATS.*JCEC.txt$",
    full.names = TRUE
)

A3SS_JCEC.path <- normalizePath(A3SS_JCEC.path)
}else{
  A3SS_JCEC.path <- list.files(
    MATS.path,
    pattern = "*A3SS.*MATS.*JCEC.txt$",
    full.names = TRUE
)

A3SS_JCEC.path <- normalizePath(A3SS_JCEC.path)
}

  
A3SS_MATS_JCEC <- tryCatch({
  read_delim(A3SS_JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))

#read in RI NovelSS
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.A3SS.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}else{
  NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.A3SS.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}

A3SS_NovelSS <- tryCatch({
  read_delim(NovelSS.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

A3SS_NovelSS <- A3SS_NovelSS %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))

#read in RI novelJunction
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.A3SS.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}else{
  NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.A3SS.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}

A3SS_NovelJunc <- tryCatch({
  read_delim(NovelJunc.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

A3SS_NovelJunc <- A3SS_NovelJunc %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))


#check if event is NovelSS
A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>%
  mutate(NovelSS = ifelse(A3SS_MATS_JCEC$SpliceID %in% A3SS_NovelSS$SpliceID, "NovelSS", ""), .after = GeneID)
#check if event is NovelJunction
A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>%
  mutate(NovelJunction = ifelse(A3SS_MATS_JCEC$SpliceID %in% A3SS_NovelJunc$SpliceID, "NovelJunction", ""), .after = NovelSS)



A3SS_MATS_JCEC <- A3SS_MATS_JCEC[order(A3SS_MATS_JCEC$FDR),]


if("A3SS_raw_rMATS_JCEC" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_raw_rMATS_JCEC")
}
writeData(wb5, "A3SS_raw_rMATS_JCEC",A3SS_MATS_JCEC )
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)

A3SS_MATS_JCEC$GeneID <- gsub("\\..*", "", A3SS_MATS_JCEC$GeneID)


# MATS_geneList <- A3SS_MATS_JCEC$geneSymbol

A3SS_MATS_JCEC <- cbind(
  A3SS_MATS_JCEC,
  with_TPM[match(A3SS_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = short_to_flanking_count)
A3SS_MATS_JCEC<- A3SS_MATS_JCEC %>%
  mutate(query = GeneID)
A3SS_MATS_JCEC = A3SS_MATS_JCEC %>%
  mutate(IGV_coords = paste0(chr, ":", longExonStart_0base, "-", flankingEE)) #finish after check w Aveksha

A3SS_MATS_JCEC <- A3SS_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("A3SS_rMATS_w_TPM" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_rMATS_w_TPM")
}
writeData(wb5, "A3SS_rMATS_w_TPM",A3SS_MATS_JCEC )
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)

JC_cutoff <- 20

SJC_ctrl <- as.vector(str_split(A3SS_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
SJC_KD <- as.vector(str_split(A3SS_MATS_JCEC$SJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})

IJC_ctrl <- as.vector(str_split(A3SS_MATS_JCEC$IJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
IJC_KD <- as.vector(str_split(A3SS_MATS_JCEC$IJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})


A3SS_COUNT_JCEC <- A3SS_MATS_JCEC[which((SJC_ctrl >= JC_cutoff | SJC_KD >= JC_cutoff) & (IJC_ctrl >= JC_cutoff | IJC_KD >= JC_cutoff)), ]

if("A3SS_JC_cutoff_20" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_JC_cutoff_20")
}
writeData(wb5, "A3SS_JC_cutoff_20", A3SS_COUNT_JCEC)
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)

#filter for padj (FDR) < 0.05
A3SS_PADJ_JCEC <- A3SS_COUNT_JCEC %>%
  filter(FDR < 0.05)

if("A3SS_padj_0.05" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_padj_0.05")
}
writeData(wb5, "A3SS_padj_0.05",A3SS_PADJ_JCEC )
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
A3SS_PSI_JCEC <- A3SS_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))

gene_info_MATS <- getGenes(geneids = A3SS_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

A3SS_PSI_JCEC <- A3SS_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

if("A3SS_dPSI_0.1" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_dPSI_0.1")
}
writeData(wb5, "A3SS_dPSI_0.1",A3SS_PSI_JCEC )
saveWorkbook(wb5, file = outfile_A3SS, overwrite = TRUE)

A3SS_PSI_annotated <- A3SS_PSI_JCEC[which((A3SS_PSI_JCEC$NovelSS) %in% "" & (A3SS_PSI_JCEC$NovelJunction) %in% ""),]

if("A3SS_filt_annotated" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_filt_annotated")
}
writeData(wb5, "A3SS_filt_annotated", A3SS_PSI_annotated)
saveWorkbook(wb5, file = outfile_A3SS, overwrite = T)

A3SS_PSI_novel <- A3SS_PSI_JCEC[which(!((A3SS_PSI_JCEC$NovelSS) %in% "") | !((A3SS_PSI_JCEC$NovelJunction) %in% "")),]

if("A3SS_filtered_novel" %in% names(wb5)){
  print("updating current worksheet")
}else{
addWorksheet(wb5, "A3SS_filtered_novel")
}
writeData(wb5, "A3SS_filtered_novel", A3SS_PSI_novel)
saveWorkbook(wb5, file = outfile_A3SS, overwrite = T)
```

#Start A5SS excel for rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_A5SS = paste0(cell_line, "_", perturbation, "_A5SS","_rMATS.xlsx")

outfile_A5SS = paste(outputdir_MATS, fileName_A5SS, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb6 <- createWorkbook()

if("README" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "README")
}
writeData(wb6, "README", README_df_M)
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)
```

#A5SS JCEC sheets
```{r}
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


A5SS_JCEC.path <- list.files(
    MATS.path,
    pattern = "*A5SS.*MATS.*JCEC.txt$",
    full.names = TRUE
)

A5SS_JCEC.path <- normalizePath(A5SS_JCEC.path)
}else{
  A5SS_JCEC.path <- list.files(
    MATS.path,
    pattern = "*A5SS.*MATS.*JCEC.txt$",
    full.names = TRUE
)

A5SS_JCEC.path <- normalizePath(A5SS_JCEC.path)
}

  
A5SS_MATS_JCEC <- tryCatch({
  read_delim(A5SS_JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})


A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))

#read in RI NovelSS
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.A5SS.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}else{
  NovelSS.path <- list.files(
    MATS.path,
    pattern = ".*novelSpliceSite.A5SS.txt$",
    full.names = TRUE
)

NovelSS.path <- normalizePath(NovelSS.path)
}

A5SS_NovelSS <- tryCatch({
  read_delim(NovelSS.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

A5SS_NovelSS <- A5SS_NovelSS %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))

#read in RI novelJunction
if(rMATS_runner == T){
MATS.path <- outputdir_MATSfiles


NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.A5SS.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}else{
  NovelJunc.path <- list.files(
    MATS.path,
    pattern = ".*novelJunction.A5SS.txt$",
    full.names = TRUE
)

NovelJunc.path <- normalizePath(NovelJunc.path)
}

A5SS_NovelJunc <- tryCatch({
  read_delim(NovelJunc.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})

A5SS_NovelJunc <- A5SS_NovelJunc %>%
  mutate(SpliceID = paste0(chr,longExonStart_0base,longExonEnd,shortES,shortEE,flankingES,flankingEE))


#check if event is NovelSS
A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>%
  mutate(NovelSS = ifelse(A5SS_MATS_JCEC$SpliceID %in% A5SS_NovelSS$SpliceID, "NovelSS", ""), .after = GeneID)
#check if event is NovelJunction
A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>%
  mutate(NovelJunction = ifelse(A5SS_MATS_JCEC$SpliceID %in% A5SS_NovelJunc$SpliceID, "NovelJunction", ""), .after = NovelSS)



A5SS_MATS_JCEC <- A5SS_MATS_JCEC[order(A5SS_MATS_JCEC$FDR),]



if("A5SS_raw_rMATS_JCEC" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_raw_rMATS_JCEC")
}
writeData(wb6, "A5SS_raw_rMATS_JCEC",A5SS_MATS_JCEC )
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)

A5SS_MATS_JCEC$GeneID <- gsub("\\..*", "", A5SS_MATS_JCEC$GeneID)


# MATS_geneList <- A5SS_MATS_JCEC$geneSymbol

A5SS_MATS_JCEC <- cbind(
  A5SS_MATS_JCEC,
  with_TPM[match(A5SS_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:(length(ctrl_bams2) + length(KD_bams2))]
)
A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = short_to_flanking_count)
A5SS_MATS_JCEC<- A5SS_MATS_JCEC %>%
  mutate(query = GeneID)
A5SS_MATS_JCEC = A5SS_MATS_JCEC %>%
  mutate(IGV_coords = paste0(chr, ":", longExonStart_0base, "-", flankingEE))

A5SS_MATS_JCEC <- A5SS_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:(length(ctrl_bams2) + length(KD_bams2))], ~ round(.x, 2)))

if("A5SS_rMATS_w_TPM" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_rMATS_w_TPM")
}
writeData(wb6, "A5SS_rMATS_w_TPM",A5SS_MATS_JCEC )
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)

JC_cutoff <- 20

SJC_ctrl <- as.vector(str_split(A5SS_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
SJC_KD <- as.vector(str_split(A5SS_MATS_JCEC$SJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})

IJC_ctrl <- as.vector(str_split(A5SS_MATS_JCEC$IJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))})
IJC_KD <- as.vector(str_split(A5SS_MATS_JCEC$IJC_SAMPLE_2, ",")) %>% sapply(function(x){mean(as.integer(x))})


A5SS_COUNT_JCEC <- A5SS_MATS_JCEC[which((SJC_ctrl >= JC_cutoff | SJC_KD >= JC_cutoff) & (IJC_ctrl >= JC_cutoff | IJC_KD >= JC_cutoff)), ]

if("A5SS_JC_cutoff_20" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_JC_cutoff_20")
}
writeData(wb6, "A5SS_JC_cutoff_20", A5SS_COUNT_JCEC)
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)

#filter for padj (FDR) < 0.05
A5SS_PADJ_JCEC <- A5SS_COUNT_JCEC %>%
  filter(FDR < 0.05)

if("A5SS_padj_0.05" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_padj_0.05")
}
writeData(wb6, "A5SS_padj_0.05",A5SS_PADJ_JCEC )
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
A5SS_PSI_JCEC <- A5SS_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))

gene_info_MATS <- getGenes(geneids = A5SS_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

A5SS_PSI_JCEC <- A5SS_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

if("A5SS_dPSI_0.1" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_dPSI_0.1")
}
writeData(wb6, "A5SS_dPSI_0.1",A5SS_PSI_JCEC )
saveWorkbook(wb6, file = outfile_A5SS, overwrite = TRUE)

A5SS_PSI_annotated <- A5SS_PSI_JCEC[which((A5SS_PSI_JCEC$NovelSS) %in% "" & (A5SS_PSI_JCEC$NovelJunction) %in% ""),]

if("A5SS_filt_annotated" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_filt_annotated")
}
writeData(wb6, "A5SS_filt_annotated", A5SS_PSI_annotated)
saveWorkbook(wb6, file = outfile_A5SS, overwrite = T)

A5SS_PSI_novel <- A5SS_PSI_JCEC[which(!((A5SS_PSI_JCEC$NovelSS) %in% "") | !((A5SS_PSI_JCEC$NovelJunction) %in% "")),]

if("A5SS_filtered_novel" %in% names(wb6)){
  print("updating current worksheet")
}else{
addWorksheet(wb6, "A5SS_filtered_novel")
}
writeData(wb6, "A5SS_filtered_novel", A5SS_PSI_novel)
saveWorkbook(wb6, file = outfile_A5SS, overwrite = T)
```

#stack annotated sheets
```{r}
SE_PSI_annotated <- SE_PSI_annotated %>%
  mutate(type = rep("SE", nrow(SE_PSI_annotated)), .before = geneSymbol)

MXE_PSI_annotated <- MXE_PSI_annotated %>%
  mutate(type = rep("MXE", nrow(MXE_PSI_annotated)), .before = geneSymbol)

RI_PSI_annotated <- RI_PSI_annotated %>%
  mutate(type = rep("RI", nrow(RI_PSI_annotated)), .before = geneSymbol)

A3SS_PSI_annotated <- A3SS_PSI_annotated %>%
  mutate(type = rep("A3SS", nrow(A3SS_PSI_annotated)), .before = geneSymbol)

A5SS_PSI_annotated <- A5SS_PSI_annotated %>%
  mutate(type = rep("A5SS", nrow(A5SS_PSI_annotated)), .before = geneSymbol)

l_list <- list(SE_PSI_annotated, MXE_PSI_annotated, RI_PSI_annotated, A3SS_PSI_annotated, A5SS_PSI_annotated)

JCEC_annotated_stacked <- rbindlist(l_list, use.names = T, fill = T)

JCEC_annotated_stacked <- as_tibble(JCEC_annotated_stacked) %>% 
  dplyr::relocate(summary, .after = short_to_flanking_count)

fileName_STACKED = paste0(cell_line, "_", perturbation, "_STACKED","_rMATS.xlsx")

outfile_STACKED = paste(outputdir_MATS, fileName_STACKED, sep = "/")


wb7 <- createWorkbook()

if("JCEC_annotated_stacked" %in% names(wb7)){
  print("updating current worksheet")
}else{
addWorksheet(wb7, "JCEC_annotated_stacked")
}
writeData(wb7, "JCEC_annotated_stacked", JCEC_annotated_stacked)
saveWorkbook(wb7, file = outfile_STACKED, overwrite = T)

```

#ggsashimi plots for top 100 events SE MATS (zoomed in)
```{r message=FALSE}
#generate ggsashimi plots for top 100 events abs(dPSI) > 0.01 and deposit to rMATS folder
library(ggplot2)
library(data.table)
library(gridExtra)

SE_top_JCEC <- SE_PSI_annotated[, c("chr", "upstreamES", "downstreamEE", "IncLevel1", "IncLevel2","IncLevelDifference", "gene", "FDR", "IJC_SAMPLE_1","SJC_SAMPLE_1", "IJC_SAMPLE_2", "SJC_SAMPLE_2")]
SE_top_JCEC <- SE_top_JCEC[order(SE_top_JCEC$IncLevelDifference),]


SE_all100_JCEC <- bind_rows(
  slice_head(SE_top_JCEC, n = 50),
  slice_tail(SE_top_JCEC, n = 50)
)

event_region <- character(nrow(SE_all100_JCEC))
event_region_expanded <- character(nrow(SE_all100_JCEC))

for(i in seq_len(nrow(SE_all100_JCEC))){
  chrom <- SE_all100_JCEC$chr[i]
  jxn_up <- SE_all100_JCEC$upstreamES[i]
  jxn_down <- SE_all100_JCEC$downstreamEE[i]
  
  jxn_leng <- jxn_down - jxn_up
  
   pading <- 300
  
  jxn_up_exp <- as.numeric(jxn_up) - pading
  jxn_down_exp <- as.numeric(jxn_down) + pading 
  
  step1 <- paste(c(chrom, jxn_up), collapse = ":")
  step2 <- paste(c(step1, jxn_down), collapse = "-")
  
  event_region[i] <- step2
  
  event_region_expanded[i] <- paste0(chrom, ":", jxn_up_exp, "-",jxn_down_exp)
  
}
SE_all100_JCEC$regions <- event_region



MATS_sashimi_regions <- event_region
MATS_exp_regions <- event_region_expanded

MATS_sashimi_genes <- SE_all100_JCEC$gene

#ggsashimi plot generation (zoomed in)

cmd <- "docker"

for(i in seq_along(MATS_exp_regions)){
  region <- MATS_exp_regions[i]
  gene_name <- MATS_sashimi_genes[i]
  
  safe_region <- gsub("[:\\-]", "_", region)
  safe_name <- paste0(i,"_",safe_region, "_", gene_name)
  print(safe_name)
  docker_args <- list(
  c("run", "--rm"),
  c("-v", paste0(normalizePath(outputdir_base), "/:/tsv")),
  c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")),
  c("-v", paste0(normalizePath(BAM_files), "/:/bams")),
  c("-v", paste0(normalizePath(outputdir_MATS_SE_allsashimi), "/:/out")),
  c("-w", "/out"),
  "guigolab/ggsashimi",
  c("-b", paste0("/tsv/", bam_tsv_name)),
  c("-c", region),
  c("-g", paste0("/gtf/", gtf_name)),
  c("-o", safe_name),
  c("-C", "3"),
  c("-M", "10"),
  "--fix-y-scale"
)


  
  tryCatch( 
    system2(cmd, args = unlist(docker_args),stdout = TRUE, stderr = TRUE),
    error = function(e){
      message("Failed region: ", region)
    }
  )
}


MATS_sashimi <- list.files(outputdir_MATS_SE_allsashimi, pattern = "\\.pdf$", full.names = TRUE)
MATS_sashimi <- MATS_sashimi[mixedorder(MATS_sashimi)]

output_file_MATS <- normalizePath(
  file.path(outputdir_MATS, "top_100_MAT_TEMP.pdf"),
  mustWork = FALSE
)
coords <- sapply(MATS_sashimi, function(coord_path){
  sub(".*/|\\.pdf$", "", coord_path)
}) %>% as.vector()


library(pdftools)
dir.create(dirname(output_file_MATS), recursive = TRUE, showWarnings = FALSE)
pdf_combine(MATS_sashimi, output = output_file_MATS)


```

#Gene-wide ggsashimi for SE rMATS (zoomed out)
```{r}
#use gtf to get gene wide coordinates for rMATS genes

gtf_wideMATS = gtf[which(gtf$feature == "gene"),]

Start = Sys.time()
MATS_gene_info <- lapply(MATS_sashimi_genes, function(gene){
  row <- gtf_wideMATS[
    grepl(paste0(".*; gene_name ", gene, ";"), gtf_wideMATS$attribute) &
    gtf_wideMATS$feature == "gene",
  ][1,]

  list(
    region = paste0(row$chr, ":", row$start, "-", row$end),
    gene_name = sub('.*gene_name ([^;]+);.*', '\\1', row$attribute)
  )
})

End = Sys.time()
print(End - Start)
MATS_regions <- sapply(MATS_gene_info, '[[', "region")
MATS_genes <- sapply(MATS_gene_info, '[[', "gene_name")


#ggsashimi generation (zoomed out--gene wide)
cmd <- "docker"

for(i in seq_along(MATS_regions)){
  region <- MATS_regions[i]
  gene_name <- MATS_genes[i]
  
  safe_region <- gsub("[:\\-]", "_", region)
  safe_name <- paste0(i, "_",safe_region, "_", gene_name)
  print(safe_name)
  docker_args <- list(
  c("run", "--rm"),
  c("-v", paste0(normalizePath(outputdir_base), "/:/tsv")),
  c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")),
  c("-v", paste0(normalizePath(BAM_files), "/:/bams")),
  c("-v", paste0(normalizePath(outputdir_MATS_SE_EXPsashimi), "/:/out")),
  c("-w", "/out"),
  "guigolab/ggsashimi",
  c("-b", paste0("/tsv/", bam_tsv_name)),
  c("-c", region),
  c("-g", paste0("/gtf/", gtf_name)),
  c("-o", safe_name),
  c("-C", "3"),
  c("-M", "5000"),
  "--fix-y-scale"
)


  
  tryCatch( 
    system2(cmd, args = unlist(docker_args),stdout = TRUE, stderr = TRUE),
    error = function(e){
      message("Failed region: ", region)
    }
  )
}

MATS_exp_sashimi <- list.files(outputdir_MATS_SE_EXPsashimi, full.names = TRUE)
MATS_exp_sashimi <- MATS_exp_sashimi[mixedorder(MATS_exp_sashimi)]

output_file_MATSexp <- normalizePath(
  file.path(outputdir_MATS, "top_100_MATexp_TEMP.pdf"),
  mustWork = FALSE
)
coords <- sapply(MATS_exp_sashimi, function(coord_path){
  sub(".*/|\\.pdf$", "", coord_path)
}) %>% as.vector()


library(pdftools)
dir.create(dirname(output_file_MATSexp), recursive = TRUE, showWarnings = FALSE)
pdf_combine(MATS_exp_sashimi, output = output_file_MATSexp)
```

```{r, message=FALSE}
#label and combine pdf with coordinates and gene name
library(pdftools)
library(png)
library(grid)

#for zoomed merged pdf
merged_pdf <- output_file_MATS
n_pages <- pdf_info(merged_pdf)$pages




tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)

annotated_pdfs <- character(n_pages)

for(i in seq_along(tmp_pngs)){

  img <- png::readPNG(tmp_pngs[i])

  out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
  annotated_pdfs[i] <- out_pdf

  pdf(out_pdf, width = 8, height = 6)

  grid.raster(img,
              x = 0.5, y = 0.5,
              width = unit(1, "npc"),
              height = unit(1, "npc"))

  grid.text(coords[i],
            x = 0.5, y = 0.97,
            gp = gpar(fontsize = 16, fontface = "bold"))

  dev.off()
}


final_pdf <- file.path(outputdir_MATS, "top_100_events_labeled.pdf")

pdf_combine(annotated_pdfs, output = final_pdf)


#for zoomed out merged pdf
merged_pdf <- output_file_MATSexp
n_pages <- pdf_info(merged_pdf)$pages




tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)

annotated_pdfs <- character(n_pages)

for(i in seq_along(tmp_pngs)){

  img <- png::readPNG(tmp_pngs[i])

  out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
  annotated_pdfs[i] <- out_pdf

  pdf(out_pdf, width = 8, height = 6)

  grid.raster(img,
              x = 0.5, y = 0.5,
              width = unit(1, "npc"),
              height = unit(1, "npc"))

  grid.text(coords[i],
            x = 0.5, y = 0.97,
            gp = gpar(fontsize = 16, fontface = "bold"))

  dev.off()
}


final_pdf <- file.path(outputdir_MATS, "top_100_EXPANDED_labeled.pdf")

pdf_combine(annotated_pdfs, output = final_pdf)
```

#MXE ggsashimi plots for top 100 events MATS (zoomed in)
```{r message=FALSE}
# #generate ggsashimi plots for top 100 events abs(dPSI) > 0.01 and deposit to rMATS folder
# library(ggplot2)
# library(data.table)
# library(gridExtra)
# 
# rm(SE_top_JCEC, SE_all100_JCEC)
# 
# SE_top_JCEC <- MXE_PSI_JCEC[, c("chr", "upstreamES", "downstreamEE", "IncLevelDifference", "gene", "FDR")]
# SE_top_JCEC <- SE_top_JCEC[order(SE_top_JCEC$IncLevelDifference),]
# 
# 
# SE_all100_JCEC <- bind_rows(
#   slice_head(SE_top_JCEC, n = 50),
#   slice_tail(SE_top_JCEC, n = 50)
# )
# 
# event_region <- character(nrow(SE_all100_JCEC))
# event_region_expanded <- character(nrow(SE_all100_JCEC))
# 
# for(i in seq_len(nrow(SE_all100_JCEC))){
#   chrom <- SE_all100_JCEC$chr[i]
#   jxn_up <- SE_all100_JCEC$upstreamES[i]
#   jxn_down <- SE_all100_JCEC$downstreamEE[i]
#   
#   jxn_leng <- jxn_down - jxn_up
#   
#    pading <- 300
#   
#   jxn_up_exp <- as.numeric(jxn_up) - pading
#   jxn_down_exp <- as.numeric(jxn_down) + pading 
#   
#   step1 <- paste(c(chrom, jxn_up), collapse = ":")
#   step2 <- paste(c(step1, jxn_down), collapse = "-")
#   
#   event_region[i] <- step2
#   
#   event_region_expanded[i] <- paste0(chrom, ":", jxn_up_exp, "-",jxn_down_exp)
#   
# }
# SE_all100_JCEC$regions <- event_region
# 
# 
# 
# MATS_sashimi_regions <- event_region
# MATS_exp_regions <- event_region_expanded
# 
# MATS_sashimi_genes <- SE_all100_JCEC$gene
# 
# #ggsashimi plot generation (zoomed in)
# 
# cmd <- "docker"
# 
# for(i in seq_along(MATS_exp_regions)){
#   region <- MATS_exp_regions[i]
#   gene_name <- MATS_sashimi_genes[i]
#   
#   safe_region <- gsub("[:\\-]", "_", region)
#   safe_name <- paste0(i,"_",safe_region, "_", gene_name)
#   print(safe_name)
#   docker_args <- list(
#   c("run", "--rm"),
#   c("-v", paste0(normalizePath(outputdir_base), "/:/tsv")),
#   c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")),
#   c("-v", paste0(normalizePath(BAM_files), "/:/bams")),
#   c("-v", paste0(normalizePath(outputdir_MATS_allsashimi), "/:/out")),
#   c("-w", "/out"),
#   "guigolab/ggsashimi",
#   c("-b", paste0("/tsv/", bam_tsv_name)),
#   c("-c", region),
#   c("-g", paste0("/gtf/", gtf_name)),
#   c("-o", safe_name),
#   c("-C", "3"),
#   c("-M", "10"),
#   "--fix-y-scale"
# )
# 
# 
#   
#   tryCatch( 
#     system2(cmd, args = unlist(docker_args),stdout = TRUE, stderr = TRUE),
#     error = function(e){
#       message("Failed region: ", region)
#     }
#   )
# }
# 
# 
# MATS_sashimi <- list.files(outputdir_MATS_allsashimi, pattern = "\\.pdf$", full.names = TRUE)
# MATS_sashimi <- MATS_sashimi[mixedorder(MATS_sashimi)]
# 
# output_file_MATS <- normalizePath(
#   file.path(outputdir_MATS, "top_100_MAT_TEMP.pdf"),
#   mustWork = FALSE
# )
# coords <- sapply(MATS_sashimi, function(coord_path){
#   sub(".*/|\\.pdf$", "", coord_path)
# }) %>% as.vector()
# 
# 
# library(pdftools)
# dir.create(dirname(output_file_MATS), recursive = TRUE, showWarnings = FALSE)
# pdf_combine(MATS_sashimi, output = output_file_MATS)
# 
# 
# ```
# 
# #MXE Gene-wide ggsashimi for rMATS (zoomed out)
# ```{r}
# #use gtf to get gene wide coordinates for rMATS genes
# 
# MATS_gene_info <- lapply(MATS_sashimi_genes, function(gene){
# 
#   row <- gtf[
#     grepl(paste0(".*; gene_name ", gene), gtf$attribute) &
#     gtf$feature == "gene",
#   ][1,]
# 
#   list(
#     region = paste0(row$chr, ":", row$start, "-", row$end),
#     gene_name = sub('.*gene_name ([^;]+);.*', '\\1', row$attribute)
#   )
# })
# 
# MATS_regions <- sapply(MATS_gene_info, '[[', "region")
# MATS_genes <- sapply(MATS_gene_info, '[[', "gene_name")
# 
# 
# #ggsashimi generation (zoomed out--gene wide)
# cmd <- "docker"
# 
# for(i in seq_along(MATS_regions)){
#   region <- MATS_regions[i]
#   gene_name <- MATS_genes[i]
#   
#   safe_region <- gsub("[:\\-]", "_", region)
#   safe_name <- paste0(i, "_",safe_region, "_", gene_name)
#   print(safe_name)
#   docker_args <- list(
#   c("run", "--rm"),
#   c("-v", paste0(normalizePath(outputdir_base), "/:/tsv")),
#   c("-v", paste0(normalizePath(gtf1_path), "/:/gtf")),
#   c("-v", paste0(normalizePath(BAM_files), "/:/bams")),
#   c("-v", paste0(normalizePath(outputdir_MATS_EXPsashimi), "/:/out")),
#   c("-w", "/out"),
#   "guigolab/ggsashimi",
#   c("-b", paste0("/tsv/", bam_tsv_name)),
#   c("-c", region),
#   c("-g", paste0("/gtf/", gtf_name)),
#   c("-o", safe_name),
#   c("-C", "3"),
#   c("-M", "5000"),
#   "--fix-y-scale"
# )
# 
# 
#   
#   tryCatch( 
#     system2(cmd, args = unlist(docker_args),stdout = TRUE, stderr = TRUE),
#     error = function(e){
#       message("Failed region: ", region)
#     }
#   )
# }
# 
# MATS_exp_sashimi <- list.files(outputdir_MATS_EXPsashimi, full.names = TRUE)
# MATS_exp_sashimi <- MATS_exp_sashimi[mixedorder(MATS_exp_sashimi)]
# 
# output_file_MATSexp <- normalizePath(
#   file.path(outputdir_MATS, "top_100_MATexp_TEMP.pdf"),
#   mustWork = FALSE
# )
# coords <- sapply(MATS_exp_sashimi, function(coord_path){
#   sub(".*/|\\.pdf$", "", coord_path)
# }) %>% as.vector()
# 
# 
# library(pdftools)
# dir.create(dirname(output_file_MATSexp), recursive = TRUE, showWarnings = FALSE)
# pdf_combine(MATS_exp_sashimi, output = output_file_MATSexp)
# ```
# 
# ```{r, message=FALSE}
# #label and combine pdf with coordinates and gene name
# library(pdftools)
# library(png)
# library(grid)
# 
# #for zoomed merged pdf
# merged_pdf <- output_file_MATS
# n_pages <- pdf_info(merged_pdf)$pages
# 
# 
# 
# 
# tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)
# 
# annotated_pdfs <- character(n_pages)
# 
# for(i in seq_along(tmp_pngs)){
# 
#   img <- png::readPNG(tmp_pngs[i])
# 
#   out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
#   annotated_pdfs[i] <- out_pdf
# 
#   pdf(out_pdf, width = 8, height = 6)
# 
#   grid.raster(img,
#               x = 0.5, y = 0.5,
#               width = unit(1, "npc"),
#               height = unit(1, "npc"))
# 
#   grid.text(coords[i],
#             x = 0.5, y = 0.97,
#             gp = gpar(fontsize = 16, fontface = "bold"))
# 
#   dev.off()
# }
# 
# 
# final_pdf <- file.path(outputdir_MATS, "top_100_events_labeled.pdf")
# 
# pdf_combine(annotated_pdfs, output = final_pdf)
# 
# 
# #for zoomed out merged pdf
# merged_pdf <- output_file_MATSexp
# n_pages <- pdf_info(merged_pdf)$pages
# 
# 
# 
# 
# tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)
# 
# annotated_pdfs <- character(n_pages)
# 
# for(i in seq_along(tmp_pngs)){
# 
#   img <- png::readPNG(tmp_pngs[i])
# 
#   out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
#   annotated_pdfs[i] <- out_pdf
# 
#   pdf(out_pdf, width = 8, height = 6)
# 
#   grid.raster(img,
#               x = 0.5, y = 0.5,
#               width = unit(1, "npc"),
#               height = unit(1, "npc"))
# 
#   grid.text(coords[i],
#             x = 0.5, y = 0.97,
#             gp = gpar(fontsize = 16, fontface = "bold"))
# 
#   dev.off()
# }
# 
# 
# final_pdf <- file.path(outputdir_MATS, "top_100_EXPANDED_labeled.pdf")
# 
# pdf_combine(annotated_pdfs, output = final_pdf)
```

#dPSI volcano plot
```{r}
library(ggplot2)
library(ggrepel)


if(!requireNamespace("pacman"))install.packages("pacman")
pacman::p_load(here,
               tidyverse,
               janitor,
               scales,
               ggrepel)

vol_plot <- SE_PSI_JCEC %>%
  ggplot(aes(x = IncLevelDifference,
             y = -log10(FDR))) +
  geom_vline(xintercept = c(0.1, -0.1),
             linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05),
             linetype = "dashed") +
  ylim(-1, 15) +
  labs(title = paste0(cell_line," ", perturbation, " dPSI distribution")) + 
  geom_point(color = "grey")+
  geom_point(data = SE_all100_JCEC,aes(x = IncLevelDifference,
             y = -log10(FDR), color = "red")) +
  geom_text_repel(data = SE_all100_JCEC,aes(x = IncLevelDifference,
             y = -log10(FDR),
             label = gene)) +
  theme_classic()
  

MATS_vol_path <- file.path(normalizePath(outputdir_MATS), "rMATS_volcano.png")
# ggsave(MATS_vol_path, plot = vol_plot, width = 8, height = 6, dpi = 300)
ggsave(MATS_vol_path, plot = vol_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)

vol_plot

```

#set up GO analysis
```{r}
library(clusterProfiler)
library(enrichplot)
library(org.Hs.eg.db)
library(org.Mm.eg.db)

JCEC_df <- SE_PSI_annotated


JCEC_df <- SE_MATS_JCEC[!is.na(SE_MATS_JCEC$FDR), ]
JCEC_df$GeneID <- sub("\\..*", "", JCEC_df$GeneID)


JCEC_sig_up <- subset(JCEC_df, FDR < 0.05 & IncLevelDifference >= 0.1)
JCEC_sig_up <- JCEC_sig_up[!is.na(JCEC_sig_up$GeneID),]
JCEC_sig_up <- JCEC_sig_up[!is.na(JCEC_sig_up$IncLevelDifference),]
JCEC_sig_up <- JCEC_sig_up$GeneID

JCEC_sig_down <- subset(JCEC_df, FDR < 0.05 & IncLevelDifference <= -0.1)
JCEC_sig_down <- JCEC_sig_down[!is.na(JCEC_sig_down$GeneID),]
JCEC_sig_down <- JCEC_sig_down[!is.na(JCEC_sig_down$IncLevelDifference),]
JCEC_sig_down <- JCEC_sig_down$GeneID

genes_sig <- JCEC_df[!is.na(JCEC_df$GeneID),]
genes_sig <- JCEC_df[!is.na(JCEC_df$FDR),]
genes_sig <- JCEC_df[!is.na(JCEC_df$IncLevelDifference),]

genes_sig <- genes_sig$GeneID

```


```{r}
#set up Webgestalt package
library(WebGestaltR)

```

```{r}
if(org == "human"){
#set up WebGestaltR function for BP
web_call_BP <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
web_call_KEGG <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
web_call_Hall <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}
}else if(org == "mouse"){
 #set up WebGestaltR function for BP
web_call_BP <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
web_call_KEGG <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
web_call_Hall <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "mmusculus",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
} 
}
```

```{r}
#GO analysis for BP (UP and DOWN)
tryCatch({
up_bp <- web_call_BP(JCEC_sig_up)
},
error = function(e){
  message("Error with GO analysis: ", e$message)
})
if(is.null(nrow(up_bp)) == T){
  print("No significant outcomes for MATS up BP")
}else{
up_bp <- web_call_BP(JCEC_sig_up)

up_bp_plot  <- ggplot(up_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (UP)", x = "enrichmentRatio", y = "Biological Process")

UpBp_path <- file.path(normalizePath(outputdir_MATS_GO), "MATS_UP_BP_GO.png")
# ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300)
ggsave(UpBp_path, plot = up_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

tryCatch({
down_bp <- web_call_BP(JCEC_sig_down)
},
error = function(e){
  message("Error with GO analysis: ", e$message)
})
if(is.null(nrow(down_bp)) == T){
  print("No significant outcomes for MATS down BP")
}else{
down_bp_plot  <- ggplot(down_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (DOWN)", x = "enrichmentRatio", y = "Biological Process")


DownBp_path <- file.path(normalizePath(outputdir_MATS_GO), "MATS_DWN_BP_GO.png")
# ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300)
ggsave(DownBp_path, plot = down_bp_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}
```

```{r}
#GO analysis for MF (UP and DOWN)
tryCatch({
up_MF <- web_call_MF(JCEC_sig_up)
}, error = function(e){
  message("Error with GO analysis: ", e$message)
})

if(is.null(nrow(up_MF)) == T){
  print("No significant outcomes for MATS up MF")
}else{
up_MF_plot  <- ggplot(up_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (UP)", x = "enrichmentRatio", y = "Molecular Function")

UpMF_path <- file.path(normalizePath(outputdir_MATS_GO), "MATS_UP_MF_GO.png")
# ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(UpMF_path, plot = up_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

tryCatch({
down_MF <- web_call_MF(JCEC_sig_down)
},
error = function(e){
  message("Error with GO analysis: ", e$message)
})

if(is.null(nrow(down_MF)) == T){
  print("No significant outcomes for MATS down MF")
}else{
down_MF_plot  <- ggplot(down_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (DOWN)", x = "enrichmentRatio", y = "Molecular Function")

DownMF_path <- file.path(normalizePath(outputdir_MATS_GO), "MATS_DWN_MF_GO.png")
# ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300)
ggsave(DownMF_path, plot = down_MF_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

```

```{r}
#GO analysis for KEGG (UP and DOWN)
tryCatch({
up_KEGG <- web_call_KEGG(JCEC_sig_up)
}, 
error = function(e){
  message("Error with GO analysis: ", e$message)
})
if(is.null(nrow(up_KEGG)) == T){
  print("No significant outcomes for MATS up KEGG")
}else{
up_KEGG_plot  <- ggplot(up_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "KEGG GO analysis (UP)", x = "enrichmentRatio", y = "KEGG")

UpKEGG_path <- file.path(normalizePath(outputdir_MATS_GO), "UP_KEGG_GO.png")
# ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 12, height = 4, dpi = 300)
ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

tryCatch({
down_KEGG <- web_call_KEGG(JCEC_sig_down)
},
error = function(e){
  message("Error with GO analysis: ", e$message)
})

if(is.null(nrow(down_KEGG)) == T){
  print("No significant outcomes for MATS down KEGG")
}else{
down_KEGG_plot  <- ggplot(down_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "KEGG GO analysis (DOWN)", x = "enrichmentRatio", y = "KEGG")

DownKEGG_path <- file.path(normalizePath(outputdir_MATS_GO), "DWN_KEGG_GO.png")
# ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 12, height = 4, dpi = 300)
ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

```

```{r}
#GO analysis for Hallmark50 (UP and DOWN)
if(org == "human"){

tryCatch({
up_Hall <- web_call_Hall(JCEC_sig_up)
},
error = function(e){
  message("Error with GO analysis: ", e$message)
})
if(is.null(nrow(up_Hall)) == T){
  print("No significant outcomes for MATS up Hallmarks 50")
}else{
up_Hall_plot  <- ggplot(up_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

UpHall_path <- file.path(normalizePath(outputdir_MATS_GO), "UP_Hallmarks_GO.png")
# ggsave(UpHall_path, plot = up_Hall_plot, width = 12, height = 4, dpi = 300)
ggsave(UpHall_path, plot = up_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

tryCatch({
down_Hall <- web_call_Hall(JCEC_sig_down)
}, 
error = function(e){
  message("Error with GO analysis: ", e$message)
})
if(is.null(nrow(down_Hall)) == T){
  print("No significant outcomes for MATS down Hallmarks 50")
}else{
down_Hall_plot  <- ggplot(down_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

DownHall_path <- file.path(normalizePath(outputdir_MATS_GO), "DWN_Hallmarks_GO.png")
# ggsave(DownHall_path, plot = down_Hall_plot, width = 12, height = 4, dpi = 300)
ggsave(DownHall_path, plot = down_Hall_plot, width = 12, height = 4, dpi = 300, 
       device = grDevices::png)
}

}else{
  message("no Hallmarks50 for mouse")
}
```

```{r}
Run_end <- "DONE"
```