---
title: "Bulk RNA-Seq DGE Pipeline"
author: "Pedro Lopez"
date: "260302"
output: html_document
editor_options: 
  chunk_output_type: inline
---

## Packages to install and use 
```{r echo=T, message=F, warning=F} 
if (!require("BiocManager", quietly = TRUE))
install.packages("BiocManager")
pkg <- c('tidyverse', "gplots", "RColorBrewer", "pheatmap", "ggplot2",
  "calibrate", "ashr", 'apeglm', "DESeq2", "dplyr", "TCGAbiolinks", "AnnotationDbi",
  "ReportingTools", "clusterProfiler", "enrichplot", "pathview", "gage", "gageData", "Select", "limma", "org.Hs.eg.db", "Rsubread", "stringr", "openxlsx", "IOBR", "mygene")

BiocManager::install(pkg, update = T, ask = FALSE, force = TRUE)

# loading
library(TCGAbiolinks)
library(dplyr)
library(DESeq2)
library(apeglm)
library(ashr)
library(calibrate)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
library(gplots)
library(tidyverse)
library(readxl)
library(AnnotationDbi)
library(ReportingTools)
library(GO.db)
library(GOstats)
library(pathview)
library(gage)
library(gageData)
library(Select)
library(limma)
library(org.Hs.eg.db)
library(Rsubread)
library(stringr)
library(openxlsx)
library(IOBR)
```
#Input relevant info
```{r}
Initials <- readline(prompt = "Enter your initials: ")

cell_line <- readline(prompt = "Enter cell line: ")

perturbation <- readline(prompt = "Enter perturbation of cell line: ")

Seq_date <- readline(prompt = "Enter date of RNAseq: ")

Replicates <- readline(prompt = "Enter number of biological replicates: ")

dir_name <- paste0(Initials, "_",cell_line, "_", perturbation)

BAM_files <- rstudioapi::selectDirectory(caption = "select folder with all BAM files", path = "/media/bree")

gtf.path <- rstudioapi::selectFile(caption = "select .gtf file used for alignment", path = "/media/bree")

#gtf.path <- "/media/bree/Expansion/gencode.v24.primary_assembly.annotation.gtf"
```

#create file directories
```{r}
outputdir <- rstudioapi::selectDirectory(caption = "Select folder for path generation")

outputdir_base <- normalizePath(file.path(outputdir, paste0("RNASeqAnalysis_",dir_name)), mustWork = FALSE)
dir.create(outputdir_base, recursive = TRUE, showWarnings = FALSE)

outputdir_SEQ <- normalizePath(file.path(outputdir_base, paste0("DESeq2_",dir_name)), mustWork = FALSE)
dir.create(outputdir_SEQ, recursive = TRUE, showWarnings = FALSE)

outputdir_SEQ_allsashimi <- normalizePath(file.path(outputdir_SEQ, "allsashimi"), mustWork = FALSE)
dir.create(outputdir_SEQ_allsashimi, recursive = TRUE, showWarnings = FALSE)

outputdir_SEQ_GO <- normalizePath(file.path(outputdir_SEQ, "GO_plots"), mustWork = FALSE)
dir.create(outputdir_SEQ_GO, recursive = TRUE, showWarnings = FALSE)

outputdir_SEQ_GSEA <- normalizePath(file.path(outputdir_SEQ, "GSEA_plots"), mustWork = FALSE)
dir.create(outputdir_SEQ_GSEA, recursive = TRUE, showWarnings = FALSE)

outputdir_MATS <-normalizePath(file.path(outputdir_base, paste0("rMATS_", dir_name)), mustWork = FALSE)
dir.create(outputdir_MATS, recursive = TRUE, showWarnings = FALSE)

outputdir_MATSfiles <- normalizePath(file.path(outputdir_MATS, "rMATS_output"), mustWork = FALSE)
dir.create(outputdir_MATSfiles, recursive = TRUE, showWarnings = FALSE)

outputdir_MATS_allsashimi <- normalizePath(file.path(outputdir_MATS, "allsashimi"), mustWork = FALSE)
dir.create(outputdir_MATS_allsashimi, recursive = TRUE, showWarnings = FALSE)

outputdir_MATS_GO <- normalizePath(file.path(outputdir_MATS, "GO_plots"), mustWork = FALSE)
dir.create(outputdir_MATS_GO, recursive = TRUE, showWarnings = FALSE)

outputdir_MATS_GSEA <- normalizePath(file.path(outputdir_MATS, "GSEA_plots"), mustWork = FALSE)
dir.create(outputdir_MATS_GSEA, recursive = TRUE, showWarnings = FALSE)

```

## generate counts for DESeq2
```{r}
bam_files <- list.files(BAM_files , full.names = TRUE)
bam_files <- gsub("\\\\", "/", bam_files)

#counting call, feed in bam files and annotations, mark as paired end

counts_mat_new <- featureCounts(files = c(unlist(bam_files[which(grepl(".bam$", bam_files)&grepl(cell_line, bam_files))])),
                                isGTFAnnotationFile = TRUE,
                                annot.ext = gtf.path, 
                                isPairedEnd = TRUE , strandSpecific = 0)

gg.bams <- c(unlist(bam_files[which(grepl("\\.bam$", bam_files)&grepl(cell_line, bam_files))]))

#make .txt files for Ctrl and KD bam paths
ctrl_bams <- c(gg.bams[which(grepl("Ctrl", gg.bams))]) %>% 
  paste(collapse = ",") %>%
  write(file.path(normalizePath(outputdir_MATS), "ctrl_bams.txt"))
ctrl_bams.path <- file.path(normalizePath(outputdir_MATS), "ctrl_bams.txt")

KD_bams <- c(gg.bams[which(!grepl("Ctrl", gg.bams))])%>%
  paste(collapse = ",") %>%
  write(file.path(normalizePath(outputdir_MATS), "KD_bams.txt"))
countdata <- counts_mat_new$counts %>%as.data.frame()
KD_bams.path <- file.path(normalizePath(outputdir_MATS), "KD_bams.txt")

#fix colnames

colname_old <- colnames(countdata)

colnames(countdata) <- sapply(colname_old, function(x){
  new_name <- str_split_1(x, "_CKD")[[1]]
  paste(new_name, "_", sep = "")
})

genes_vec <- rownames(countdata)
countdata <- mutate(countdata, Gene = genes_vec, .before = colnames(countdata)[1])

#copy and paste your spreadsheet into read_excel(___). 



fileName = paste0(cell_line, "_", perturbation, "_DESeq2.xlsx")

outfile = paste(outputdir_SEQ, fileName, sep = "/")

#README sheet at the start of excel
Sheet <- c("Overall description", "", "", "counts matrix", "", "", "DESeq2 output", "", "",
                       "FDR_0.05", "", "", "|log2FoldChange|_1","","","","","","")
t_ <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description <- c("List all output data associated with DESeq2 pipeline", "", "", "raw read counts used for DESeq2 run", "", "", "unfiltered DESeq2 output. Includes log2FoldChange; log2foldchange standard error (lfcSE); p-value and p adj value (padj or FDR); ensembl ID; gene name (SYMBOL); TPM; Description of gene", "", "", "filtered for padjusted values less than 0.05", "", "", "filtered FDR_0.05 filtered data set further to include only genes with an absolute log2FoldChange greater than 1", "", "", Initials, cell_line, perturbation, date())
README_df <- data.frame(Sheet = Sheet, t_ = t_, Sheet_description = Sheet_description )

wb <- createWorkbook()
addWorksheet(wb, "README")
writeData(wb, "README", README_df)
addWorksheet(wb, "counts matrix")
writeData(wb, "counts matrix", countdata)
saveWorkbook(wb, file = outfile, overwrite = TRUE)
```

## Prepare data for Analysis with DESeq2
```{r}
## Upload raw read count matrix as a table
counts_mat <- read_excel(outfile, sheet = "counts matrix")

## Convert the table to a dataframe

counts_mat<-as.data.frame(counts_mat)

## Assign the gene symbol/ID to row names, remove the gene symbol/ID column, 
# asummes that there is a column called Gene
row.names(counts_mat) <- counts_mat$Gene
counts_mat <- counts_mat[ ,-which(colnames(counts_mat) %in% 'Gene')]


## Assign condition and other variables (Ensure the order of the factor matches
## the position/column of the sample within the raw read count matrix)

Group <- factor(c(rep("Ctrl",Replicates), rep("KD",Replicates)))
# Batch <- factor(c("Batch1","Batch2","Batch1","Batch2")) <- removed due to RNASeq protocol mod to remove                                                               batch effect 260310


## Create the colData variable (table of sample information) for DESeq2 
## based on the conditions 
coldata<- data.frame(row.names=colnames(counts_mat),Group)
```

## Differential Gene Expression Analysis with DESeq2
```{r}
## Generate a DESEQDataSet File from the matrix, column data, and group variables

dds <- DESeqDataSetFromMatrix(countData = counts_mat,
                              colData = coldata,
                              design = ~Group)

## Run the DESeq function

dds <- DESeq(dds)

## Generate Dispersion Plot for QC purposes

Disp_path <- file.path(normalizePath(outputdir_SEQ), "qc-dispersions.pdf")
pdf(Disp_path, width = 7, height=5)
plotDispEsts(dds, main="Dispersion plot")
dev.off()

## Perform the differential gene expression analysis and acquire the results 
## between the two conditions (use ashr method to generate log2FoldChange 
## values)

res <- results(dds, alpha = 0.05)
res <- lfcShrink(dds, contrast = c("Group","KD","Ctrl"),
type = "ashr")


## Generate a MA plot (scatterplot of the log fold change and mean average of
## normalized counts)

MA_path <- file.path(normalizePath(outputdir_SEQ), "MAplot.png")
png(MA_path, 1000, 1000, pointsize=20)
plotMA(res,ylim=c(-5,5))
dev.off()


## Generate volcano plot

"Gene" <- names(res)[1]
volcanoplot <- function (res, lfcthresh=1, sigthresh=0.05, main="Volcano Plot",
legendpos="bottomright", labelsig=TRUE, textcx=1, ...)
{
with(res, plot(log2FoldChange, -log10(pvalue), pch=20, main=main, ...))
with(subset(res, padj<sigthresh ), points(log2FoldChange, -log10(pvalue),
pch=20, col="red", ...))
with(subset(res, abs(log2FoldChange)>lfcthresh), points(log2FoldChange,
-log10(pvalue), pch=20, col="orange", ...))
with(subset(res, padj<sigthresh & abs(log2FoldChange)>lfcthresh),
points(log2FoldChange, -log10(pvalue), pch=20, col="green", ...))
if (labelsig) {
require(calibrate)
4
with(subset(res, padj<sigthresh & abs(log2FoldChange)>lfcthresh),
textxy(log2FoldChange, -log10(pvalue), labs=NULL, cex=textcx, ...))
}
legend(legendpos, xjust=1, yjust=1, legend=c(paste("FDR<",sigthresh,sep=""),
paste("|LogFC|>",lfcthresh,sep=""), "both"),
pch=20, col=c("red","orange","green"))
}
vol_path <- file.path(normalizePath(outputdir_SEQ), "volcano.png")
png(vol_path, 1200, 1000, pointsize=20)
volcano <- volcanoplot(res, lfcthresh=1, sigthresh=0.05, textcx=.8)
dev.off()

## Perform a variance stabilizing transformation on the DESeq data and create matrix
## with vst transformed count data

vstdata <- vst(dds)
vstMatrix <- assay(vstdata)

##Generate Sample-Sample Distance Heatmap with vst() function

sampleDists <- as.dist(1-cor(vstMatrix))
sampleDistMatrix <- as.matrix(sampleDists)
annoCol <- as.data.frame(colData(dds))
annoCol=annoCol[c(1)]

heat_path <- file.path(normalizePath(outputdir_SEQ), "heatmap.png")
heat_map <- pheatmap(sampleDistMatrix,
clustering_distance_rows=sampleDists,
clustering_distance_cols=sampleDists ,annotation_col = annoCol,
filename = heat_path,
width = 8,
height = 10)
heat_map

```

## OPTIONAL Alternative Sample Distance Plot
```{r}
# Transpose count matrix and calculate distances using dist()
sampleDists <- assay(vstdata) %>% t() %>% dist()

# Convert distance dataframe to matrix
sampleDistMatrix <- as.matrix(sampleDists)

# Choose continuous palette from RColorBrewer
colors <- colorRampPalette(rev(brewer.pal(9, "Blues")))(255)

# Plot sample distance heatmap with ComplexHeatmap
altHeat_path <- file.path(normalizePath(outputdir_SEQ), "Alt_sampleDistance_plot.png")
pheatmap(sampleDistMatrix, cluster_rows = T, cluster_columns = T,
        col = colors,main= cell_line,
        filename = altHeat_path)
```


## Generate PCAplot
```{r}
library(ggplot2)
rld = rlog(dds)
PCA <- plotPCA(rld, intgroup = "Group", ntop = 2000)
pca_path <- file.path(normalizePath(outputdir_SEQ), "PCA_plot.png")
ggsave(pca_path, plot = PCA, width = 8, height = 6, dpi = 300)
PCA
```

## Generate log2 normalized gene expression scatterplot 
```{r}
## generate dataframe of normalized expression values for a gene of interest

SFPQexpression <- plotCounts(dds = dds,gene = "ENSG00000116560.10",intgroup = "Group",returnData = TRUE) #need to utomate gene = step 

##create a new dataframe and convert the normalized counts to log2 normalized 
##counts for visualization purposes

logSFPQexpression <-SFPQexpression
logSFPQexpression$count<-log2(logSFPQexpression$count)

##create the plot 
l2f <- ggplot(logSFPQexpression , aes(x=Group, y=count)) + ylab(paste0("Log2 Normalized ",  cell_line," Expression")) + xlab("Group") +  geom_point(shape=1) + ggtitle(cell_line) + theme(plot.title = element_text(hjust = 0.5))+geom_text(aes(label = rownames(logSFPQexpression)),hjust = -0.2, vjust = 0.5)
l2f_path <- file.path(normalizePath(outputdir_SEQ), "l2f_norm_gene_expresScatterplot.png")
ggsave(l2f_path, plot = l2f, width = 8, height = 6, dpi = 300)
l2f
```


## Generate Heatmap of all Significant DE Genes with p-value <0.05 and
## L2FC < -1 (Sig. Downregulated Genes in KD)
```{r}
sigChanges <- rownames(res)[res$padj < 0.05 & res$log2FoldChange < -1 ]
sigMat <- vstMatrix[rownames(vstMatrix) %in% sigChanges,]
sigHDWN <- pheatmap(sigMat, scale="row", show_rownames = FALSE,show_colnames = FALSE,
col=greenred(75), border_color = NA, annotation_col = annoCol)
sigHeatDWN_path <- file.path(normalizePath(outputdir_SEQ), "HeatsigGene_down.png")
ggsave(sigHeatDWN_path, plot = sigHDWN, width = 8, height = 6, dpi = 300)
sigHDWN
```

## Generate Heatmap of all Significant DE Genes with p-value <0.05 and
## L2FC > 1 (Sig. Upregulated Genes in KD)
```{r}
sigChanges <- rownames(res)[res$padj < 0.05 & res$log2FoldChange > 1 ]
sigMat <- vstMatrix[rownames(vstMatrix) %in% sigChanges,]
sigHUP <- pheatmap(sigMat, scale="row", show_rownames = FALSE,show_colnames = FALSE,
col=greenred(75), border_color = NA, annotation_col = annoCol)
sigHeatUP_path <- file.path(normalizePath(outputdir_SEQ), "HeatsigGene_up.png")
ggsave(sigHeatUP_path, plot = sigHUP, width = 8, height = 6, dpi = 300)
sigHUP
```
#print out list of genes that are siginificantly different from Ctrl to KD

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
```

```{r}
res$ENSEMBL = gsub("\\..*","",row.names(res))
 

```

```{r}
res$gene = mapIds(org.Hs.eg.db,
                    keys = res$ENSEMBL,
                    column = "SYMBOL",
                    keytype = "ENSEMBL",
                    multiVals = "first")

res$ENTREZ = mapIds(org.Hs.eg.db,
                    keys = res$ENSEMBL,
                    column = "ENTREZID",
                    keytype = "ENSEMBL",
                    multiVals = "first")
head(res)
```

#Generate TPM data frame from counts matrix
```{r}
rownames(counts_mat) <- gsub("\\..*", "", rownames(counts_mat))
with_TPM  <-IOBR::count2tpm(counts_mat, idType = "Ensembl", org = "hsa")
with_TPM$gene <- rownames(with_TPM)



res_df <- as.data.frame(res)

library(mygene)

res_df<- res_df[!is.na(res_df$padj), ] #check with Dr. Cheng if na's should be removed
res_df <- res_df[order(res_df$padj),]

gene_info <- getGenes(geneids = res_df$ENSEMBL, fields = "summary") 
gene_deduped <- gene_info[which(!duplicated(gene_info$query)),]
gene_desc <- as.vector(gene_deduped$summary)


res_geneList <- res_df$gene

res_df <- cbind(
  res_df,
  with_TPM[match(res_df$gene, rownames(with_TPM)),1:6]
)
res_df <- res_df %>%
  mutate(Gene = res_df$gene, .before = 1)
res_df <- res_df %>% 
  mutate(gene = res_df$gene)
res_df <- res_df%>%
  mutate(Description = gene_desc)
res_df <- res_df %>%
  mutate(across(colnames(with_TPM)[1:6], ~ round(.x, 2)))

addWorksheet(wb, "DESeq2 output")
writeData(wb, "DESeq2 output", res_df)
saveWorkbook(wb, file = outfile, overwrite = TRUE)
```

```{r}
#make sheet with filters for cutoffs of DESEq2 output
res_df_padj <- res_df %>%
  filter(padj < 0.05)
addWorksheet(wb, "FDR_0.05")
writeData(wb, "FDR_0.05", res_df_padj)
saveWorkbook(wb, file = outfile, overwrite = TRUE)

res_df_allfilt <- res_df_padj %>%
  filter(abs(log2FoldChange) >= 1 )
addWorksheet(wb, "|log2FoldChange| _ 1")
writeData(wb, "|log2FoldChange| _ 1", res_df_allfilt)
saveWorkbook(wb, file = outfile, overwrite = TRUE)
```

##DESeq Gene Ontology Analysis
```{r}
library(clusterProfiler)
library(WebGestaltR)
```

```{r}
sigUp <- subset(res_df, padj < 0.05 & log2FoldChange >=1)
sigUp <- sigUp[-which(is.na(sigUp$ENTREZ)),]
genes_sig_Up <- sigUp$ENTREZ 

sigDown <- subset(res_df, padj < 0.05 & log2FoldChange <= -1 )
sigDown <- sigDown[-which(is.na(sigDown$ENTREZ)),]
genes_sig_down <- sigDown$ENTREZ

sigAll <- res_df[-which(is.na(res_df$ENTREZ)),]
sigAll <- sigAll$ENTREZ

```


```{r}
universeGenes = res_df$gene

```

```{r}
#set up WebGestaltR function for BP
web_call_BP <- function(x, cutoff = 0.05, bkg = sigAll, id = "entrezgene",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = sigAll, id = "entrezgene", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
web_call_KEGG <- function(x, cutoff = 0.05, bkg = sigAll, id = "entrezgene", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
web_call_Hall <- function(x, cutoff = 0.05, bkg = sigAll, id = "entrezgene", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

```

```{r}
#GO analysis for BP (UP and DOWN)

up_bp <- web_call_BP(genes_sig_Up)

up_bp_plot  <- ggplot(up_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (UP)", x = "enrichmentRatio", y = "Biological Process")

down_bp <- web_call_BP(genes_sig_down)

down_bp_plot  <- ggplot(down_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (DOWN)", x = "enrichmentRatio", y = "Biological Process")

UpBp_path <- file.path(normalizePath(outputdir_SEQ), "UP_biological_processes_GO.png")
DownBp_path <- file.path(normalizePath(outputdir_SEQ), "DWN_biological_processes_GO.png")

ggsave(UpBp_path, plot = up_bp_plot, width = 8, height = 3, dpi = 300)
ggsave(DownBp_path, plot = down_bp_plot, width = 8, height = 3, dpi = 300)

up_bp_plot
down_bp_plot
```

```{r}
#GO analysis for MF (UP and DOWN)

up_MF <- web_call_MF(genes_sig_Up)

up_MF_plot  <- ggplot(up_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (UP)", x = "enrichmentRatio", y = "Molecular Function")


down_MF <- web_call_MF(genes_sig_down)

down_MF_plot  <- ggplot(down_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (DOWN)", x = "enrichmentRatio", y = "Molecular Function")

UpMF_path <- file.path(normalizePath(outputdir_SEQ), "UP_molecular_function_GO.png")
DownMF_path <- file.path(normalizePath(outputdir_SEQ), "DWN_molecular_function_GO.png")

ggsave(UpMF_path, plot = up_MF_plot, width = 6, height = 3, dpi = 300)
ggsave(DownMF_path, plot = down_MF_plot, width = 6, height = 3, dpi = 300)

up_MF_plot
down_MF_plot
```

```{r}
#GO analysis for KEGG (UP and DOWN)

up_KEGG <- web_call_KEGG(genes_sig_Up)

up_KEGG_plot  <- ggplot(up_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "KEGG GO analysis (UP)", x = "enrichmentRatio", y = "KEGG")

down_KEGG <- web_call_KEGG(genes_sig_down)

down_KEGG_plot  <- ggplot(down_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "KEGG GO analysis (DOWN)", x = "enrichmentRatio", y = "KEGG")

UpKEGG_path <- file.path(normalizePath(outputdir_SEQ), "UP_KEGG_GO.png")
DownKEGG_path <- file.path(normalizePath(outputdir_SEQ), "DWN_KEGG_GO.png")

ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 6, height = 3, dpi = 300)
ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 6, height = 3, dpi = 300)

up_KEGG_plot
down_KEGG_plot
```

```{r}
#GO analysis for Hallmark50 (UP and DOWN)

up_Hall <- web_call_Hall(genes_sig_Up)

up_Hall_plot  <- ggplot(up_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

down_Hall <- web_call_Hall(genes_sig_down)

down_Hall_plot  <- ggplot(down_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

UpHall_path <- file.path(normalizePath(outputdir_SEQ), "UP_Hallmarks_GO.png")
DownHall_path <- file.path(normalizePath(outputdir_SEQ), "DWN_Hallmarks_GO.png")

ggsave(UpHall_path, plot = up_Hall_plot, width = 6, height = 3, dpi = 300)
ggsave(DownHall_path, plot = down_Hall_plot, width = 6, height = 3, dpi = 300)

up_Hall_plot
down_Hall_plot
```

#GSEA analysis
```{r}
library(DESeq2)
library(org.Hs.eg.db)
library(AnnotationDbi)

res_raw <- DESeq2::results(dds)

res_raw$ENSEMBL <- gsub("\\..*", "", rownames(res_raw))

res_raw$ENTREZ <- mapIds(
  org.Hs.eg.db,
  keys = res_raw$ENSEMBL,
  column = "ENTREZID",
  keytype = "ENSEMBL",
  multiVals = "first"
)

res_raw <- res_raw[!is.na(res_raw$ENTREZ) & !is.na(res_raw$log2FoldChange), ]

gene_list <- res_raw$log2FoldChange
names(gene_list) <- as.character(res_raw$ENTREZ)
gene_list <- sort(gene_list, decreasing = TRUE)

gene_df <- data.frame(
  gene = names(gene_list),
  score = gene_list
)

```

```{r}
#set up WebGestaltR function for BP
GSEA_call_BP <- function(x, cutoff = 0.05, id = "entrezgene",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = TRUE,
    isOutput = TRUE
  )
}

#WebGestaltR function for MF
GSEA_call_MF <- function(x, cutoff = 0.05, id = "entrezgene", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
GSEA_call_KEGG <- function(x, cutoff = 0.05, id = "entrezgene", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,


    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
GSEA_call_Hall <- function(x, cutoff = 0.05, id = "entrezgene", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "GSEA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

  
    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = F,
    isOutput = F
)
}

```

```{r}
GSEA_BP <- GSEA_call_BP(gene_df)

GSEA_MF <- GSEA_call_MF(gene_df)

GSEA_KEGG <- GSEA_call_KEGG(gene_df)

GSEA_Hall <- GSEA_call_Hall(gene_df)


GSEABP_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_BP.png")
GSEAMF_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_MF.png")
GSEAKEGG_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_KEGG.png")
GSEAHall_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_Hallmarks.png")

```

```{r}
GSEA_BP_plot <- ggplot(GSEA_BP, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Biological Processes",
    x = "Normalized Enrichment Score",
    y = "Biological Processes",
    size = "Gene count",
    color = "FDR"
  )


GSEA_MF_plot <- ggplot(GSEA_MF, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Molecular Function",
    x = "Normalized Enrichment Score",
    y = "Molecular Function",
    size = "Gene count",
    color = "FDR"
  )

GSEA_KEGG_plot <- ggplot(GSEA_KEGG, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA KEGG",
    x = "Normalized Enrichment Score",
    y = "KEGG",
    size = "Gene count",
    color = "FDR"
  )

GSEA_Hall_plot <- ggplot(GSEA_Hall, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) +
  geom_point()+
  scale_color_gradient(
    low = "blue", high = "red"
  ) +
  scale_size(range = c(3,10)) +
  theme_bw() +
  labs(
    title = "GSEA Hallmarks 50",
    x = "Normalized Enrichment Score",
    y = "Hallmarks",
    size = "Gene count",
    color = "FDR"
  )
ggsave(GSEABP_path, plot = GSEA_BP, width = 4, height = 3, dpi = 300)
ggsave(GSEAMF_path, plot = GSEA_MF, width = 8, height = 6, dpi = 300)
ggsave(GSEAKEGG_path, plot = GSEA_KEGG, width = 8, height = 6, dpi = 300)
ggsave(GSEAHall_path, plot = GSEA_Hall, width = 8, height = 6, dpi = 300)


GSEA_BP_plot
GSEA_MF_plot
GSEA_KEGG_plot
GSEA_Hall_plot
```

```{r}

library(fgsea)
library(ggplot2)

output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_BP.pdf")

pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches

top_sets <- GSEA_BP[1:10, ]

for (i in seq_len(nrow(top_sets))) {
    
    geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]]
    geneset <- intersect(geneset, names(gene_list))
    
    if (length(geneset) < 5) {
        message("Skipping: ", top_sets$description[i])
        next
    }
    
    # Create GSEA plot
    p <- plotEnrichment(geneset, gene_list) +
        ggtitle(paste0(
            top_sets$description[i],
            "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2),
            ", FDR = ", signif(top_sets$FDR[i], 2)
        ))
    
    # Print to PDF (each print() = new page)
    print(p)
}

dev.off()

message("PDF saved to: ", output_seqGSEA)

#MF GSEA
output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_MF.pdf")

pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches

top_sets <- GSEA_MF[1:10, ]

for (i in seq_len(nrow(top_sets))) {
    
    geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]]
    geneset <- intersect(geneset, names(gene_list))
    
    if (length(geneset) < 5) {
        message("Skipping: ", top_sets$description[i])
        next
    }
    
    # Create GSEA plot
    p <- plotEnrichment(geneset, gene_list) +
        ggtitle(paste0(
            top_sets$description[i],
            "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2),
            ", FDR = ", signif(top_sets$FDR[i], 2)
        ))
    
    # Print to PDF (each print() = new page)
    print(p)
}

dev.off()

message("PDF saved to: ", output_seqGSEA)

#KEGG GSEA
output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_KEGG.pdf")

pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches

top_sets <- GSEA_KEGG[1:10, ]

for (i in seq_len(nrow(top_sets))) {
    
    geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]]
    geneset <- intersect(geneset, names(gene_list))
    
    if (length(geneset) < 5) {
        message("Skipping: ", top_sets$description[i])
        next
    }
    
    # Create GSEA plot
    p <- plotEnrichment(geneset, gene_list) +
        ggtitle(paste0(
            top_sets$description[i],
            "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2),
            ", FDR = ", signif(top_sets$FDR[i], 2)
        ))
    
    # Print to PDF (each print() = new page)
    print(p)
}

dev.off()

message("PDF saved to: ", output_seqGSEA)

#Hallmarks50 GSEA
output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_Hall.pdf")

pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches

top_sets <- GSEA_Hall[1:10, ]

for (i in seq_len(nrow(top_sets))) {
    
    geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]]
    geneset <- intersect(geneset, names(gene_list))
    
    if (length(geneset) < 5) {
        message("Skipping: ", top_sets$description[i])
        next
    }
    
    # Create GSEA plot
    p <- plotEnrichment(geneset, gene_list) +
        ggtitle(paste0(
            top_sets$description[i],
            "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2),
            ", FDR = ", signif(top_sets$FDR[i], 2)
        ))
    
    # Print to PDF (each print() = new page)
    print(p)
}

dev.off()

message("PDF saved to: ", output_seqGSEA)
```


#ggsashimi DESEq2 top 100 events
```{r}
#generate ggsashimi plots for top 100 events abs(dPSI) > 0.01 and deposit to rMATS folder
library(ggplot2)
library(data.table)
library(gridExtra)

#generate .tsv table for ggsashimi DESEq run
SEQ_top100 <- bind_rows(
  slice_head(res_df_allfilt, n = 100)
)
res_ggsashimi_genes <- rownames(SEQ_top100) 

gtf <- read.delim(
  gtf.path,
  header = FALSE,
  comment.char = "#",
  stringsAsFactors = FALSE
)

colnames(gtf) <- c(
  "chr","source","feature","start","end",
  "score","strand","frame","attribute"
)

gene_info <- lapply(res_ggsashimi_genes, function(gene){

  row <- gtf[
    grepl(paste0("gene_id ", gene), gtf$attribute) &
    gtf$feature == "gene",
  ][1,]

  list(
    region = paste0(row$chr, ":", row$start, "-", row$end),
    gene_name = sub('.*gene_name ([^;]+);.*', '\\1', row$attribute)
  )
})

DEseq_regions <- sapply(gene_info, '[[', "region")
Seq_genes <- sapply(gene_info, '[[', "gene_name")


colors_key <- c(rep("#1f77b4",Replicates), rep("#ff7f0e", Replicates))
#gg.bams <- c(unlist(bam_files[which(grepl("\\.bam$", bam_files)&grepl("_LM2", bam_files))]))
reps <- 1:Replicates
rep_nums <- rep(reps, each = 1)
Ungrp <- unfactor(Group) %>% paste0(rep_nums)
bams.tsv <-  data.frame(group = Ungrp, bams = gg.bams, colorsK = colors_key )

write_tsv(bams.tsv, file.path(outputdir_base, "gg_bams.tsv"))



ggsashimi_path <- path.expand("~/ggsashimi/ggsashimi.py")
bam_tsv <- normalizePath(file.path(outputdir_base, "gg_bams.tsv"))

  
for(i in seq_along(DEseq_regions)){
  
  region <- DEseq_regions[i]
  gene_name <- Seq_genes[i]

  safe_region <- gsub("[:\\-]", "_", region)
  safe_name <- paste0(safe_region, "_", gene_name)
  
  
  tryCatch(
  system2("/home/bree/miniforge3/envs/rmats_env/bin/python3", args = c(
    ggsashimi_path,
    "-b", shQuote(bam_tsv),
    "-c", shQuote(region),
    "-C", 3,
    "-M", 5000,
    "--fix-y-scale",
    "-g", shQuote("/media/bree/Expansion/gencode.v24.primary_assembly.annotation.gtf"),
    "-o", shQuote(file.path(outputdir_SEQ_allsashimi, paste0(safe_name, ".pdf")))
    )
  ),
    error = function(e){
      message("Failed region: ", region)
    } 
 
  )
  
}


DESEQ_sashimi <- list.files(outputdir_SEQ_allsashimi, pattern = "\\.pdf$", full.names = TRUE)
#order of gene names for labeled merged pdf
genes <- sapply(DESEQ_sashimi, function(name_path){
  sub('.*\\d_([^_]+)\\..*', "\\1", name_path)
}) %>% as.vector()

output_file <- normalizePath(
  file.path(outputdir_SEQ, "temp_100Events.pdf"),
  mustWork = FALSE
)


library(pdftools)
dir.create(dirname(output_file), recursive = TRUE, showWarnings = FALSE)
pdf_combine(DESEQ_sashimi, output = output_file)

library(pdftools)
library(grid)
library(grDevices)

merged_pdf <- output_file

n_pages <- pdf_info(merged_pdf)$pages

```

#label and merge Seq sashimi plots for final pdf
```{r}
library(pdftools)
library(png)
library(grid)

merged_pdf <- output_file
n_pages <- pdf_info(merged_pdf)$pages


tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)

annotated_pdfs <- character(n_pages)

for(i in seq_along(tmp_pngs)){

  img <- png::readPNG(tmp_pngs[i])

  out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
  annotated_pdfs[i] <- out_pdf

  pdf(out_pdf, width = 8, height = 6)

  grid.raster(img,
              x = 0.5, y = 0.5,
              width = unit(1, "npc"),
              height = unit(1, "npc"))

  grid.text(genes[i],
            x = 0.5, y = 0.97,
            gp = gpar(fontsize = 16, fontface = "bold"))

  dev.off()
}


final_pdf <- file.path(outputdir_SEQ, "top_100_events_labeled.pdf")

pdf_combine(annotated_pdfs, output = final_pdf)
```



#Run rMATS in terminal
```{r}
rMATS_path <- path.expand("~/rmats_env/bin/python/rmats.py")
system2("/home/bree/miniforge3/envs/rmats_env/bin/python", args = c(
  rMATS_path,
  "--b1", ctrl_bams.path,
  "--b2", KD_bams.path,
  "--gtf", gtf.path,
  "-t", "paired",
  "--readLength", 150,
  "--nthread", 12,
  "--od", outputdir_MATSfiles,
  "--tmp", outputdir_MATSfiles,
  "--individual-counts", 
  "--novelSS",
  "--variable-read-length"
  )
  
)
```

#Start excel for rMATS outputs
```{r}
#select folder to send rMATS data processing output


fileName_M = paste0(cell_line, "_", perturbation, "_rMATS.xlsx")

outfile_M = paste(outputdir_MATS, fileName_M, sep = "/")

#README sheet at the start of excel
Sheet_M <- c("Overall description", "", "", "raw_rMATS_JCEC", "", "", "rMATS_w_TPM", "", "",
                       "padj_0.05", "", "", "dPSI_0.1", "", "", "", "", "", "")
t_M <- c("-->", "", "", "-->", "", "", "-->", "", "",
                       "-->", "", "", "-->","","", "Initials","Cell line", "Perturbation", "Date of processing")
Sheet_description_M <- c("List all output data associated with rMATS pipeline", "", "", "raw rMATS output for JCEC (junction reads and exon reads) file", "", "", "rMATS JCEC output with TPM according to ensembl ID. For more information regarding column names refer to link below:", "https://github.com/Xinglab/rmats-turbo#output", "", "dPSI_0.1 data filtered further for padj values (FDR) < 0.05", "", "", "rMATS_w_TPM data filtered for dPSI abs(IncLevelDifference) > 0.1", "", "", Initials, cell_line, perturbation, date())
README_df_M <- data.frame(Sheet = Sheet_M, t_ = t_M, Sheet_description = Sheet_description_M )

wb2 <- createWorkbook()
addWorksheet(wb2, "README")
writeData(wb2, "README", README_df_M)
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)
```


```{r}
#Raw rMATS output
MATS.path <- outputdir_MATSfiles

MATS_files <- list.files(MATS.path, full.names = TRUE)

JCEC.path <- list.files(
    "/media/bree/Main_storage/siSFPQ_LM2_rMATS_output",
    pattern = "*SE.*MATS.*JCEC.txt$",
    full.names = TRUE
)

JCEC.path <- normalizePath(JCEC.path)



  
SE_MATS_JCEC <- tryCatch({
  read_delim(JCEC.path, delim = "\t", na = c("", "NA"))
}, error = function(e){
  message("Error reading file: ", e$message)
  NULL
})
SE_MATS_JCEC <- SE_MATS_JCEC[order(SE_MATS_JCEC$FDR),]
addWorksheet(wb2, "raw_rMATS_JCEC")
writeData(wb2, "raw_rMATS_JCEC",SE_MATS_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

SE_MATS_JCEC$GeneID <- gsub("\\..*", "", SE_MATS_JCEC$GeneID)


MATS_geneList <- SE_MATS_JCEC$geneSymbol

SE_MATS_JCEC <- cbind(
  SE_MATS_JCEC,
  with_TPM[match(SE_MATS_JCEC$geneSymbol, rownames(with_TPM)),1:6]
)
SE_MATS_JCEC <- SE_MATS_JCEC %>%
  relocate(geneSymbol, .before = 1)
SE_MATS_JCEC <- SE_MATS_JCEC %>% 
  mutate(gene = geneSymbol, .after = upstream_to_downstream_count)
SE_MATS_JCEC<- SE_MATS_JCEC %>%
  mutate(query = GeneID)

SE_MATS_JCEC <- SE_MATS_JCEC %>%
  mutate(across(colnames(with_TPM)[1:6], ~ round(.x, 2)))

addWorksheet(wb2, "rMATS_w_TPM")
writeData(wb2, "rMATS_w_TPM",SE_MATS_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

SJC_inidividual <- as.vector(str_split(SE_MATS_JCEC$SJC_SAMPLE_1, ",")) %>% sapply(function(x){mean(as.integer(x))}) 

SE_COUNT_JCEC <- SE_MATS_JCEC[which(SJC_inidividual >= 10), ]

addWorksheet(wb2, "ctrl_SJC_10")
writeData(wb2, "ctrl_SJC_10", SE_COUNT_JCEC)
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

#filter for padj (FDR) < 0.05
SE_PADJ_JCEC <- SE_COUNT_JCEC %>%
  filter(FDR < 0.05)
addWorksheet(wb2, "padj_0.05")
writeData(wb2, "padj_0.05",SE_PADJ_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

#filter for dPSI and remove NAs from Inc groups 1 and 2  
SE_PSI_JCEC <- SE_PADJ_JCEC %>% 
  filter(abs(IncLevelDifference) > 0.1) %>% 
  filter(!grepl("NA", IncLevel1)) %>% filter(!grepl("NA", IncLevel2))

gene_info_MATS <- getGenes(geneids = SE_PSI_JCEC$GeneID, fields = "summary") 
gene_deduped_MATS <- gene_info_MATS[which(!duplicated(gene_info_MATS$query)),]
gene_desc_MATS <- gene_deduped_MATS[,c("query", "summary")] %>% as.data.frame()

SE_PSI_JCEC <- SE_PSI_JCEC %>%
  left_join(gene_desc_MATS, by = "query")

addWorksheet(wb2, "dPSI_0.1")
writeData(wb2, "dPSI_0.1",SE_PSI_JCEC )
saveWorkbook(wb2, file = outfile_M, overwrite = TRUE)

```
#ggsashimi plots for top 100 events MATS
```{r}
#generate ggsashimi plots for top 100 events abs(dPSI) > 0.01 and deposit to rMATS folder
library(ggplot2)
library(data.table)
library(gridExtra)

SE_top_JCEC <- SE_PSI_JCEC[, c("chr", "upstreamEE", "downstreamES", "IncLevelDifference", "gene", "FDR")]
SE_top_JCEC <- SE_top_JCEC[order(SE_top_JCEC$IncLevelDifference),]
SE_top100_JCEC <- bind_rows(
  slice_head(SE_top_JCEC, n = 100)
)

SE_bottom100_JCEC <- bind_rows(
  slice_tail(SE_top_JCEC, n = 100)
)

event_regionstop <- character(nrow(SE_top100_JCEC))
event_regionTOP_expanded <- character(nrow(SE_top100_JCEC))

for(i in seq_len(nrow(SE_top100_JCEC))){
  chrom <- SE_top100_JCEC$chr[i]
  jxn_up <- SE_top100_JCEC$upstreamEE[i]
  jxn_down <- SE_top100_JCEC$downstreamES[i]
  
  jxn_leng <- jxn_down - jxn_up
  
   pading <- max(300, min(1000, round(jxn_leng * 0.5)))
  
  jxn_up_exp <- max(1, as.numeric(jxn_up) - pading)
  jxn_down_exp <- as.numeric(jxn_down) + pading 
  
  step1 <- paste(c(chrom, jxn_up), collapse = ":")
  step2 <- paste(c(step1, jxn_down), collapse = "-")
  
  event_regionstop[i] <- step2
  
  event_regionTOP_expanded[i] <- paste0(chrom, ":", jxn_up_exp, "-",jxn_down_exp)
  
}
SE_top100_JCEC$regions <- event_regionstop

event_regionsbot <- character(nrow(SE_bottom100_JCEC))
event_regionBOT_expanded <- character(nrow(SE_bottom100_JCEC))
for(i in seq_len(nrow(SE_bottom100_JCEC))){
  chrom <- SE_bottom100_JCEC$chr[i]
  jxn_up <- SE_bottom100_JCEC$upstreamEE[i]
  jxn_down <- SE_bottom100_JCEC$downstreamES[i]
  
  jxn_leng <- jxn_down - jxn_up
  
  pading <- max(300, min(1000, round(jxn_leng * 0.5)))
  
  jxn_up_exp <- max(1, as.numeric(jxn_up) - pading)
  jxn_down_exp <- as.numeric(jxn_down) + pading 
  
  step1 <- paste(c(chrom, jxn_up), collapse = ":")
  step2 <- paste(c(step1, jxn_down), collapse = "-")
  
  event_regionsbot[i] <- step2
  
  event_regionBOT_expanded[i] <- paste0(chrom, ":", jxn_up_exp, "-",jxn_down_exp)
  
  
}
SE_bottom100_JCEC$regions <- event_regionsbot

MATSTOP_sashimi_regions <- event_regionstop
MATSBOT_sashimi_regions <- event_regionsbot

MATSTOP_sashimi_genes <- SE_top100_JCEC$gene
MATSBOT_sashimi_genes <- SE_bottom100_JCEC$gene

#top 100
for(i in seq_along(MATSTOP_sashimi_regions)){
  
  region <- event_regionTOP_expanded[i]
  gene_name <- MATSTOP_sashimi_genes[i]
  
  safe_region <- gsub("[:\\-]", "_", i)
  safe_name <- paste0(safe_region, "_", gene_name, "_top100")
  
  tryCatch(
  system2("/home/bree/miniforge3/envs/rmats_env/bin/python3", args = c(
    ggsashimi_path,
    "-b", shQuote(bam_tsv),
    "-c", shQuote(region),
    "-L", 1,
    "-C", 3,
    "-g", shQuote(gtf.path),
    "-o", shQuote(file.path(outputdir_MATS_allsashimi, paste0(safe_name, ".pdf")))
    )
  ), 
   error = function(e){ message("Failed region: ", i)}
  )
  
}

#bottom 100
for(i in seq_along(MATSBOT_sashimi_regions)){
  
  region <- event_regionBOT_expanded[i]
  gene_name <- MATSBOT_sashimi_genes[i]
  
  safe_region <- gsub("[:\\-]", "_", i)
  safe_name <- paste0(safe_region, "_", gene_name,"_bottom100")
  
  tryCatch(
  system2("/home/bree/miniforge3/envs/rmats_env/bin/python3", args = c(
    ggsashimi_path,
    "-b", shQuote(bam_tsv),
    "-c", shQuote(region),
    "-L", 1,
    "-C", 3,
    "--width", 25,
    "--shrink",
    "-g", shQuote(gtf.path),
    "-o", shQuote(file.path(outputdir_MATS_allsashimi, paste0(safe_name, ".pdf")))
    )
  ), 
   error = function(e){ message("Failed region: ", i)}
  )
  
}

MATS_sashimi <- list.files(outputdir_MATS_allsashimi, pattern = "\\.pdf$", full.names = TRUE)

output_file_MATS <- normalizePath(
  file.path(outputdir_MATS, "top_100_MAT_TEMP.pdf"),
  mustWork = FALSE
)
coords <- sapply(MATS_sashimi, function(coord_path){
  sub(".*/|\\.pdf$", "", coord_path)
}) %>% as.vector()


library(pdftools)
dir.create(dirname(output_file_MATS), recursive = TRUE, showWarnings = FALSE)
pdf_combine(MATS_sashimi, output = output_file_MATS)


```

```{r}
#label and combine pdf with coordinates and gene name
library(pdftools)
library(png)
library(grid)

merged_pdf <- output_file_MATS
n_pages <- pdf_info(merged_pdf)$pages




tmp_pngs <- pdf_convert(merged_pdf, format = "png", dpi = 150)

annotated_pdfs <- character(n_pages)

for(i in seq_along(tmp_pngs)){

  img <- png::readPNG(tmp_pngs[i])

  out_pdf <- file.path(tempdir(), sprintf("page_%03d.pdf", i))
  annotated_pdfs[i] <- out_pdf

  pdf(out_pdf, width = 8, height = 6)

  grid.raster(img,
              x = 0.5, y = 0.5,
              width = unit(1, "npc"),
              height = unit(1, "npc"))

  grid.text(coords[i],
            x = 0.5, y = 0.97,
            gp = gpar(fontsize = 16, fontface = "bold"))

  dev.off()
}


final_pdf <- file.path(outputdir_MATS, "top_100_events_labeled.pdf")

pdf_combine(annotated_pdfs, output = final_pdf)
```

#dPSI volcano plot
```{r}
library(ggplot2)
library(ggrepel)


if(!requireNamespace("pacman"))install.packages("pacman")
pacman::p_load(here,
               tidyverse,
               janitor,
               scales,
               ggrepel)

vol_plot <- SE_PSI_JCEC %>%
  ggplot(aes(x = IncLevelDifference,
             y = -log10(FDR))) +
  geom_vline(xintercept = c(0.1, -0.1),
             linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05),
             linetype = "dashed") +
  ylim(-1, 15) +
  labs(title = paste0(cell_line," ", perturbation, " dPSI distribution")) + 
  geom_point(color = "grey")+
  geom_point(data = SE_top100_JCEC,aes(x = IncLevelDifference,
             y = -log10(FDR), color = "red")) +
  geom_text_repel(data = SE_top100_JCEC,aes(x = IncLevelDifference,
             y = -log10(FDR),
             label = gene)) +
  theme_classic()
  

MATS_vol_path <- file.path(normalizePath(outputdir_MATS), "rMATS_volcano.png")
ggsave(MATS_vol_path, plot = vol_plot, width = 8, height = 6, dpi = 300)

vol_plot

```

#set up GO analysis
```{r}
library(clusterProfiler)
library(enrichplot)
library(org.Hs.eg.db)

JCEC_df <- SE_MATS_JCEC

JCEC_df <- SE_PSI_JCEC

# JCEC_df <- SE_MATS_JCEC[!is.na(SE_MATS_JCEC$FDR), ]
# JCEC_df$GeneID <- sub("\\..*", "", JCEC_df$GeneID)

JCEC_sig_up <- subset(JCEC_df, FDR < 0.05 & IncLevelDifference >= 0.1)
JCEC_sig_up <- JCEC_sig_up[!is.na(JCEC_sig_up$GeneID),]
JCEC_sig_up <- JCEC_sig_up[!is.na(JCEC_sig_up$IncLevelDifference),]
JCEC_sig_up <- JCEC_sig_up$GeneID %>% unique()

JCEC_sig_down <- subset(JCEC_df, FDR < 0.05 & IncLevelDifference <= -0.1)
JCEC_sig_down <- JCEC_sig_down[!is.na(JCEC_sig_down$GeneID),]
JCEC_sig_down <- JCEC_sig_down[!is.na(JCEC_sig_down$IncLevelDifference),]
JCEC_sig_down <- JCEC_sig_down$GeneID %>% unique()

genes_sig <-JCEC_df$GeneID %>% unique()

```


```{r}
#set up Webgestalt package
library(WebGestaltR)

```

```{r}
#set up WebGestaltR function for BP
web_call_BP <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id",
                       path = 'geneontology_Biological_Process_noRedundant'){

  WebGestaltR(
    enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
  )
}

#WebGestaltR function for MF
web_call_MF <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for KEGG
web_call_KEGG <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'pathway_KEGG'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

#WebGestaltR function for Hallmark50
web_call_Hall <- function(x, cutoff = 0.05, bkg = genes_sig, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){
WebGestaltR(
   enrichMethod = "ORA",
    organism = "hsapiens",
    enrichDatabase = path,

    interestGene = x,
    interestGeneType = id,

    referenceGene = bkg,
    referenceGeneType = id,

    minNum = 5,
    maxNum = 500,
    sigMethod = "fdr",
    fdrMethod = "BH",
    fdrThr = cutoff,
    topThr = 10000,

    saveRawGseaResult = FALSE,
    isOutput = FALSE
)
}

```

```{r}
#GO analysis for BP (UP and DOWN)

up_bp <- web_call_BP(JCEC_sig_up, cutoff = 0.25)

up_bp <- web_call_BP(JCEC_sig_up, cutoff = 0.75)


up_bp_plot  <- ggplot(up_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (UP)", x = "enrichmentRatio", y = "Biological Process")

down_bp <- web_call_BP(JCEC_sig_down, cutoff = 0.5)

down_bp_plot  <- ggplot(down_bp)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(title = "Biological Processes GO analysis (DOWN)", x = "enrichmentRatio", y = "Biological Process")

UpBp_path <- file.path(normalizePath(outputdir_MATS), "MATS_UP_BP_GO.png")
DownBp_path <- file.path(normalizePath(outputdir_MATS), "MATS_DWN_BP_GO.png")

ggsave(UpBp_path, plot = up_bp_plot, width = 8, height = 3, dpi = 300)
ggsave(DownBp_path, plot = down_bp_plot, width = 8, height = 3, dpi = 300)

up_bp_plot
down_bp_plot
```

```{r}
#GO analysis for MF (UP and DOWN)

up_MF <- web_call_MF(JCEC_sig_up, cutoff = 1)

up_MF_plot  <- ggplot(up_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (UP)", x = "enrichmentRatio", y = "Molecular Function")


down_MF <- web_call_MF(JCEC_sig_down)

down_MF_plot  <- ggplot(down_MF)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Molecular Function GO analysis (DOWN)", x = "enrichmentRatio", y = "Molecular Function")

UpMF_path <- file.path(normalizePath(outputdir_MATS), "MATS_UP_MF_GO.png")
DownMF_path <- file.path(normalizePath(outputdir_MATS), "MATS_DWN_MF_GO.png")

ggsave(UpMF_path, plot = up_MF_plot, width = 6, height = 3, dpi = 300)
ggsave(DownMF_path, plot = down_MF_plot, width = 6, height = 3, dpi = 300)

up_MF_plot
down_MF_plot
```

```{r}
#GO analysis for KEGG (UP and DOWN)

up_KEGG <- web_call_KEGG(JCEC_sig_up, cutoff = 0.75)

up_KEGG_plot  <- ggplot(up_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "KEGG GO analysis (UP)", x = "enrichmentRatio", y = "KEGG")

down_KEGG <- web_call_KEGG(JCEC_sig_down)

down_KEGG_plot  <- ggplot(down_KEGG)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "KEGG GO analysis (DOWN)", x = "enrichmentRatio", y = "KEGG")

UpKEGG_path <- file.path(normalizePath(outputdir_SEQ), "UP_KEGG_GO.png")
DownKEGG_path <- file.path(normalizePath(outputdir_SEQ), "DWN_KEGG_GO.png")

ggsave(UpKEGG_path, plot = up_KEGG_plot, width = 6, height = 3, dpi = 300)
ggsave(DownKEGG_path, plot = down_KEGG_plot, width = 6, height = 3, dpi = 300)

up_KEGG_plot
down_KEGG_plot
```

```{r}
#GO analysis for Hallmark50 (UP and DOWN)

up_Hall <- web_call_Hall(JCEC_sig_up, cutoff = 0.5)

up_Hall_plot  <- ggplot(up_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "blue") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

down_Hall <- web_call_Hall(JCEC_sig_down)

down_Hall_plot  <- ggplot(down_Hall)+
  geom_col(aes(x = enrichmentRatio, y = reorder(description, enrichmentRatio)), fill = "orange") +
  theme_classic() +
  labs(title = "Hallmark50 GO analysis (UP)", x = "enrichmentRatio", y = "Hallmarks 50")

UpHall_path <- file.path(normalizePath(outputdir_SEQ), "UP_Hallmarks_GO.png")
DownHall_path <- file.path(normalizePath(outputdir_SEQ), "DWN_Hallmarks_GO.png")

ggsave(UpHall_path, plot = up_Hall_plot, width = 6, height = 3, dpi = 300)
ggsave(DownHall_path, plot = down_Hall_plot, width = 6, height = 3, dpi = 300)

up_Hall_plot
down_Hall_plot
```
#commented out until GSEA method is fixed for rMATS output
<!-- #GSEA analysis -->
<!-- ```{r} -->
<!-- library(DESeq2) -->
<!-- library(org.Hs.eg.db) -->
<!-- library(AnnotationDbi) -->

<!-- res_raw <- DESeq2::results(dds) -->

<!-- res_raw$ENSEMBL <- gsub("\\..*", "", rownames(res_raw)) -->

<!-- res_raw$ENTREZ <- mapIds( -->
<!--   org.Hs.eg.db, -->
<!--   keys = res_raw$ENSEMBL, -->
<!--   column = "ENTREZID", -->
<!--   keytype = "ENSEMBL", -->
<!--   multiVals = "first" -->
<!-- ) -->

<!-- res_raw <- res_raw[!is.na(res_raw$ENTREZ) & !is.na(res_raw$log2FoldChange), ] -->

<!-- gene_list <- res_raw$log2FoldChange -->
<!-- names(gene_list) <- as.character(res_raw$ENTREZ) -->
<!-- gene_list <- sort(gene_list, decreasing = TRUE) -->

<!-- gene_df <- data.frame( -->
<!--   gene = names(gene_list), -->
<!--   score = gene_list -->
<!-- ) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- #set up WebGestaltR function for BP -->
<!-- GSEA_call_BP <- function(x, cutoff = 0.05, id = "ensembl_gene_id", -->
<!--                        path = 'geneontology_Biological_Process_noRedundant'){ -->

<!--   WebGestaltR( -->
<!--     enrichMethod = "GSEA", -->
<!--     organism = "hsapiens", -->
<!--     enrichDatabase = path, -->

<!--     interestGene = x, -->
<!--     interestGeneType = id, -->

<!--     minNum = 5, -->
<!--     maxNum = 500, -->
<!--     sigMethod = "fdr", -->
<!--     fdrMethod = "BH", -->
<!--     fdrThr = cutoff, -->
<!--     topThr = 10000, -->

<!--     saveRawGseaResult = TRUE, -->
<!--     isOutput = TRUE -->
<!--   ) -->
<!-- } -->

<!-- #WebGestaltR function for MF -->
<!-- GSEA_call_MF <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'geneontology_Molecular_Function_noRedundant'){ -->
<!-- WebGestaltR( -->
<!--    enrichMethod = "GSEA", -->
<!--     organism = "hsapiens", -->
<!--     enrichDatabase = path, -->

<!--     interestGene = x, -->
<!--     interestGeneType = id, -->

<!--     minNum = 5, -->
<!--     maxNum = 500, -->
<!--     sigMethod = "fdr", -->
<!--     fdrMethod = "BH", -->
<!--     fdrThr = cutoff, -->
<!--     topThr = 10000, -->

<!--     saveRawGseaResult = FALSE, -->
<!--     isOutput = FALSE -->
<!-- ) -->
<!-- } -->

<!-- #WebGestaltR function for KEGG -->
<!-- GSEA_call_KEGG <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'pathway_KEGG'){ -->
<!-- WebGestaltR( -->
<!--    enrichMethod = "GSEA", -->
<!--     organism = "hsapiens", -->
<!--     enrichDatabase = path, -->

<!--     interestGene = x, -->
<!--     interestGeneType = id, -->


<!--     minNum = 5, -->
<!--     maxNum = 500, -->
<!--     sigMethod = "fdr", -->
<!--     fdrMethod = "BH", -->
<!--     fdrThr = cutoff, -->
<!--     topThr = 10000, -->

<!--     saveRawGseaResult = FALSE, -->
<!--     isOutput = FALSE -->
<!-- ) -->
<!-- } -->

<!-- #WebGestaltR function for Hallmark50 -->
<!-- GSEA_call_Hall <- function(x, cutoff = 0.05, id = "ensembl_gene_id", path = 'community-contributed_Hallmark50'){ -->
<!-- WebGestaltR( -->
<!--    enrichMethod = "GSEA", -->
<!--     organism = "hsapiens", -->
<!--     enrichDatabase = path, -->

<!--     interestGene = x, -->
<!--     interestGeneType = id, -->


<!--     minNum = 5, -->
<!--     maxNum = 500, -->
<!--     sigMethod = "fdr", -->
<!--     fdrMethod = "BH", -->
<!--     fdrThr = cutoff, -->
<!--     topThr = 10000, -->

<!--     saveRawGseaResult = F, -->
<!--     isOutput = F -->
<!-- ) -->
<!-- } -->

<!-- ``` -->

<!-- ```{r} -->
<!-- GSEA_BP <- GSEA_call_BP(gene_df) -->

<!-- GSEA_MF <- GSEA_call_MF(gene_df) -->

<!-- GSEA_KEGG <- GSEA_call_KEGG(gene_df) -->

<!-- GSEA_Hall <- GSEA_call_Hall(gene_df) -->


<!-- GSEABP_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_BP.png") -->
<!-- GSEAMF_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_MF.png") -->
<!-- GSEAKEGG_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_KEGG.png") -->
<!-- GSEAHall_path <- file.path(normalizePath(outputdir_SEQ), "GSEA_Hallmarks.png") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- GSEA_BP_plot <- ggplot(GSEA_BP, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) + -->
<!--   geom_point()+ -->
<!--   scale_color_gradient( -->
<!--     low = "blue", high = "red" -->
<!--   ) + -->
<!--   scale_size(range = c(3,10)) + -->
<!--   theme_bw() + -->
<!--   labs( -->
<!--     title = "GSEA Biological Processes", -->
<!--     x = "Normalized Enrichment Score", -->
<!--     y = "Biological Processes", -->
<!--     size = "Gene count", -->
<!--     color = "FDR" -->
<!--   ) -->


<!-- GSEA_MF_plot <- ggplot(GSEA_MF, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) + -->
<!--   geom_point()+ -->
<!--   scale_color_gradient( -->
<!--     low = "blue", high = "red" -->
<!--   ) + -->
<!--   scale_size(range = c(3,10)) + -->
<!--   theme_bw() + -->
<!--   labs( -->
<!--     title = "GSEA Molecular Function", -->
<!--     x = "Normalized Enrichment Score", -->
<!--     y = "Molecular Function", -->
<!--     size = "Gene count", -->
<!--     color = "FDR" -->
<!--   ) -->

<!-- GSEA_KEGG_plot <- ggplot(GSEA_KEGG, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) + -->
<!--   geom_point()+ -->
<!--   scale_color_gradient( -->
<!--     low = "blue", high = "red" -->
<!--   ) + -->
<!--   scale_size(range = c(3,10)) + -->
<!--   theme_bw() + -->
<!--   labs( -->
<!--     title = "GSEA KEGG", -->
<!--     x = "Normalized Enrichment Score", -->
<!--     y = "KEGG", -->
<!--     size = "Gene count", -->
<!--     color = "FDR" -->
<!--   ) -->

<!-- GSEA_Hall_plot <- ggplot(GSEA_Hall, aes(x = normalizedEnrichmentScore, y = reorder(description, normalizedEnrichmentScore), size = size, color = -log10(FDR))) + -->
<!--   geom_point()+ -->
<!--   scale_color_gradient( -->
<!--     low = "blue", high = "red" -->
<!--   ) + -->
<!--   scale_size(range = c(3,10)) + -->
<!--   theme_bw() + -->
<!--   labs( -->
<!--     title = "GSEA Hallmarks 50", -->
<!--     x = "Normalized Enrichment Score", -->
<!--     y = "Hallmarks", -->
<!--     size = "Gene count", -->
<!--     color = "FDR" -->
<!--   ) -->
<!-- ggsave(GSEABP_path, plot = GSEA_BP, width = 4, height = 3, dpi = 300) -->
<!-- ggsave(GSEAMF_path, plot = GSEA_MF, width = 8, height = 6, dpi = 300) -->
<!-- ggsave(GSEAKEGG_path, plot = GSEA_KEGG, width = 8, height = 6, dpi = 300) -->
<!-- ggsave(GSEAHall_path, plot = GSEA_Hall, width = 8, height = 6, dpi = 300) -->


<!-- GSEA_BP_plot -->
<!-- GSEA_MF_plot -->
<!-- GSEA_KEGG_plot -->
<!-- GSEA_Hall_plot -->
<!-- ``` -->
<!-- ```{r} -->

<!-- library(fgsea) -->
<!-- library(ggplot2) -->

<!-- output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_BP.pdf") -->

<!-- pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches -->

<!-- top_sets <- GSEA_BP[1:10, ] -->

<!-- for (i in seq_len(nrow(top_sets))) { -->

<!--     geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]] -->
<!--     geneset <- intersect(geneset, names(gene_list)) -->

<!--     if (length(geneset) < 5) { -->
<!--         message("Skipping: ", top_sets$description[i]) -->
<!--         next -->
<!--     } -->

<!--     # Create GSEA plot -->
<!--     p <- plotEnrichment(geneset, gene_list) + -->
<!--         ggtitle(paste0( -->
<!--             top_sets$description[i], -->
<!--             "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2), -->
<!--             ", FDR = ", signif(top_sets$FDR[i], 2) -->
<!--         )) -->

<!--     # Print to PDF (each print() = new page) -->
<!--     print(p) -->
<!-- } -->

<!-- dev.off() -->

<!-- message("PDF saved to: ", output_seqGSEA) -->

<!-- #MF GSEA -->
<!-- output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_MF.pdf") -->

<!-- pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches -->

<!-- top_sets <- GSEA_MF[1:10, ] -->

<!-- for (i in seq_len(nrow(top_sets))) { -->

<!--     geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]] -->
<!--     geneset <- intersect(geneset, names(gene_list)) -->

<!--     if (length(geneset) < 5) { -->
<!--         message("Skipping: ", top_sets$description[i]) -->
<!--         next -->
<!--     } -->

<!--     # Create GSEA plot -->
<!--     p <- plotEnrichment(geneset, gene_list) + -->
<!--         ggtitle(paste0( -->
<!--             top_sets$description[i], -->
<!--             "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2), -->
<!--             ", FDR = ", signif(top_sets$FDR[i], 2) -->
<!--         )) -->

<!--     # Print to PDF (each print() = new page) -->
<!--     print(p) -->
<!-- } -->

<!-- dev.off() -->

<!-- message("PDF saved to: ", output_seqGSEA) -->

<!-- #KEGG GSEA -->
<!-- output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_KEGG.pdf") -->

<!-- pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches -->

<!-- top_sets <- GSEA_KEGG[1:10, ] -->

<!-- for (i in seq_len(nrow(top_sets))) { -->

<!--     geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]] -->
<!--     geneset <- intersect(geneset, names(gene_list)) -->

<!--     if (length(geneset) < 5) { -->
<!--         message("Skipping: ", top_sets$description[i]) -->
<!--         next -->
<!--     } -->

<!--     # Create GSEA plot -->
<!--     p <- plotEnrichment(geneset, gene_list) + -->
<!--         ggtitle(paste0( -->
<!--             top_sets$description[i], -->
<!--             "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2), -->
<!--             ", FDR = ", signif(top_sets$FDR[i], 2) -->
<!--         )) -->

<!--     # Print to PDF (each print() = new page) -->
<!--     print(p) -->
<!-- } -->

<!-- dev.off() -->

<!-- message("PDF saved to: ", output_seqGSEA) -->

<!-- #Hallmarks50 GSEA -->
<!-- output_seqGSEA <- file.path(outputdir_SEQ_GSEA, "GSEA_Hall.pdf") -->

<!-- pdf(output_seqGSEA, width = 6, height = 5)  # width/height in inches -->

<!-- top_sets <- GSEA_Hall[1:10, ] -->

<!-- for (i in seq_len(nrow(top_sets))) { -->

<!--     geneset <- strsplit(top_sets$leadingEdgeId[i], ";")[[1]] -->
<!--     geneset <- intersect(geneset, names(gene_list)) -->

<!--     if (length(geneset) < 5) { -->
<!--         message("Skipping: ", top_sets$description[i]) -->
<!--         next -->
<!--     } -->

<!--     # Create GSEA plot -->
<!--     p <- plotEnrichment(geneset, gene_list) + -->
<!--         ggtitle(paste0( -->
<!--             top_sets$description[i], -->
<!--             "\nNES = ", round(top_sets$normalizedEnrichmentScore[i], 2), -->
<!--             ", FDR = ", signif(top_sets$FDR[i], 2) -->
<!--         )) -->

<!--     # Print to PDF (each print() = new page) -->
<!--     print(p) -->
<!-- } -->

<!-- dev.off() -->

<!-- message("PDF saved to: ", output_seqGSEA) -->
<!-- ``` -->
